---
title: 'Protein turnover analysis in mouse and human cells'
author:
    - name: "Henrik Hammarén, PhD"
      email: "henrik.hammaren@gmail.com, henrik.m.hammaren@gsk.com"
date: "`r Sys.Date()`"
output: 
  rmdformats::html_clean:
    code_folding: hide
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango
editor_options:
  chunk_output_type: inline   
description: >
---

# Set up

```{r setup}
knitr::opts_chunk$set(echo = TRUE,
                      include = T,
                      message = F,
                      warning = F)

VERSION <- "v05"

FONT_SIZE_PAPER <- 8
PLOT_TEXT_SIZE <- 2

OUTLIER_SIZE <- 0.3
LINEWIDTH <- 0.3

```

## Load packages

```{r Load packages, warning=F, message=F}

library(tidyverse)
library(readr)
library(beeswarm)
library(ggrepel)
library(gridExtra)
library(grid)
library(ggbeeswarm)
library(Hmisc)
library(ggpubr)
library(pheatmap)
library(RColorBrewer)
library(UpSetR)
library(GGally)
library(viridis)
library(MASS) 
library(plotly)
library(BiocManager)
library(magrittr)
library(devtools)
library(eulerr)
library(ggridges)
library(pheatmap)
library(cowplot)
library(stringr)
library(svglite)
library(fgsea)
library(ggpointdensity)
library(fabricatr)
library(scales)
library(xlsx)


if (!require(vsn)) BiocManager::install("vsn")
library(vsn)

if (!require(limma)) BiocManager::install("limma")
library(limma)
 
if (!require(org.Hs.eg.db)) BiocManager::install("org.Hs.eg.db")
library("org.Hs.eg.db")
# 
if (!require(org.Mm.eg.db)) BiocManager::install("org.Mm.eg.db")
library("org.Mm.eg.db")
# 
if (!require(GO.db)) BiocManager::install("GO.db")
library("GO.db")
# 
if (!require(biomaRt)) BiocManager::install("biomaRt")
library("biomaRt")

if (!require(clusterProfiler)) BiocManager::install("clusterProfiler")
library("clusterProfiler")

```

## Set functions

```{r Functions, warning=F, message=F}

COLOR_VECTOR_SAMPLE <- c('human' = 'orange',
                         'mouse' = 'steelblue',
                         'mouse_2DG' = '#c08ced',
                         'mouse\n2DG' = '#c08ced')

## Set colors for subcellular annotation
COLOR_CC <- c("extracellular matrix" = "#91638e",
              "extracellular region" = "#DD8DD8",
              "plasma membrane" = "brown",
              "cytoskeleton" = "#D53E4F",
              "cytoplasm" = "#FC8D59",
              "cytosol" = "#FEE08B",
              "mitochondrion" = "#66D775",
              "lysosome" = "#57f2de",
              "Golgi apparatus" = "#73bab1",
              "peroxisome" = "#74cca0",
              "ER" = "#2f8a7e",
              "nuclear membrane" = "#8b9af7",
              "nucleus" = "#6A78C8",
              "nucleoplasm" = "#8791cc",
              "nucleolus" = "#253699",
              "ambiguous" = "grey")

COLOR_HITS <- c("faster in human" = "orange",
                "slower in human" = "steelblue",
                "no difference" = "grey")

COLOR_HITS_2DG <- c("faster in 2DG" = "#c08ced",
                "slower in 2DG" = "steelblue",
                "no difference" = "grey")

theme_HH_paper <- list(theme_classic(base_size = FONT_SIZE_PAPER),
                         theme(
                           line = element_line(linewidth = 0.3),
                           panel.background = element_rect(fill = "transparent"),
                           panel.grid.major.x = element_blank(),
                           #panel.grid.major.y = element_line(linetype = "dashed", color = "grey80"),
                           panel.grid.major.y = element_blank(),
                           axis.line.x = element_line(),
                           axis.title = element_text(size = FONT_SIZE_PAPER),
                           axis.text = element_text(size = FONT_SIZE_PAPER, color = "black"),
                           legend.text = element_text(size = FONT_SIZE_PAPER, color = "black"),
                           title = element_text(size = FONT_SIZE_PAPER),
                           strip.background = element_blank()
                           ),
                         guides(colour = guide_legend(override.aes = list(alpha = 1,
                                                                          size = 1))))

theme_HH_scatter_paper <- list(theme_classic(base_size = FONT_SIZE_PAPER),
                         theme(
                           line = element_line(linewidth = 0.3),
                           panel.grid.major.x = element_line(linewidth = 0.1, linetype = "solid", color = "grey90"),
                           panel.grid.major.y = element_line(linewidth = 0.1, linetype = "solid", color = "grey90"),
                           panel.grid.minor.x = element_line(linewidth = 0.1, linetype = "solid", color = "grey90"),
                           panel.grid.minor.y = element_line(linewidth = 0.1, linetype = "solid", color = "grey90"),
                           axis.line.x = element_blank(),
                           axis.line.y = element_blank(),
                           axis.ticks = element_line(linewidth = 0.3),
                           axis.title = element_text(size = FONT_SIZE_PAPER),
                           axis.text = element_text(size = FONT_SIZE_PAPER, color = "black"),
                           legend.text = element_text(size = FONT_SIZE_PAPER, color = "black"),
                           title = element_text(size = FONT_SIZE_PAPER),
                           strip.background = element_blank(),
                           panel.background = element_blank(),
                           panel.border = element_rect(linewidth = 0.3, color = "black",
                                                       fill = NA),
                           plot.background = element_blank(),
                           plot.margin = grid::unit(c(0,0,0,0), "mm")
                           ),
                         guides(colour = guide_legend(override.aes = list(alpha = 1,
                                                                          size = 1))))


l_workflow <- list()

## Make function to extract number of entries at different levels of the workflow
f_n_entries <- function(fdf, stage = "") {
  c("stage" = stage,
    "n_gene_name" = n_distinct(fdf$gene_name))
}

```

## Read in data

```{r Read in data, warning = FALSE}
#  Read in metadata
metadata <- read_delim(list.files(pattern = "metadata_HH.csv$"), ",")

## Order metadata
metadata <- metadata[order(metadata$time),]

# list all txt files from the current directory
list.filenames<-list.files(path = "../Data/",
                           pattern=".txt$")

# Check that files in metadata and folder match 
if(all(metadata$source_file[order(metadata$source_file)] == list.filenames[order(list.filenames)])) {
  print("All files in metadata file found.", quote = 0)
} else {
  print("Files in metadata and folder do not match.", quote = 0)
  knitr::knit_exit()
}

# create an empty list that will serve as a container to receive the incoming files
list.data<-list()

# create a loop to read in and analyze data
for (i in 1:length(list.filenames)) {
  
  #Read in data
  list.data[[i]]<-read_delim(paste0("../Data/", list.filenames[i]), "\t")
  
  print(paste0("Reading in ", list.filenames[i]), quote = 0)
  
  ## Add info on, where data came from
  list.data[[i]]$source_file <- list.filenames[i]
  
  ## Add column with total signal sum info
  list.data[[i]]$Total_Signal_Sum <- rowSums(list.data[[i]][,grep("signal_sum", colnames(list.data[[i]]))])
}

# Combine all dataframes into one
df_raw_data <- do.call(rbind, list.data)
df <- df_raw_data

rm(i, list.data, dups, before, after, list.filenames)

```

## Filtering of data (remove contaminants, etc., filter for qupm, etc.)

```{r Clean and prepare data}

  #Remove rows with contaminants
  before <- n_distinct(df$gene_name)
  df <- df[!grepl("###",df$protein_id),] 
  print(paste0("Removed ", before-n_distinct(df$gene_name), " contaminant or reverse (decoy) gene names."))
  
  #Remove entries with less than 2 qupm
  before <- n_distinct(df$gene_name)
  df <- filter(df, qupm > 1)
  print(paste0("Started with ", before, " non-contaminant gene names across both species./nRemoved ", before-n_distinct(df$gene_name), " gene names with qupm < 2."))

  #Remove rows without quantification data
  before <- n_distinct(df$gene_name)
  df <- df[df$Total_Signal_Sum > 0,]
  df <- df[!is.na(df$Total_Signal_Sum), ]
  print(paste0("Removed ", before-n_distinct(df$gene_name), " gene names with no quantification data at all."))

  #Remove rows without H2L data
  before <- n_distinct(df$gene_name)
  df <- df[!grepl("NaN", df$rel_fc_HEAVY),]
  df <- df[!grepl("Inf", df$rel_fc_HEAVY),]
  df <- df[!is.na(df$rel_fc_HEAVY),]
  df <- df[df$rel_fc_HEAVY > 0,]
  print(paste0("Removed ", before-n_distinct(df$gene_name), " gene names with no SILAC ratios (as at least one SILAC channel was missing)"))

  print(paste0("Left with ", n_distinct(df$gene_name), " gene names across both species over all input data."))

  rm(before, after)
```

```{r}
rm(df_raw_data)
```

### Adding metadata

```{r Add metadata}

df <- left_join(df,
                metadata %>% dplyr::select(-Fraction),
                by = c("source_file")) %>% 
    mutate(Sample = factor(Sample,
                         levels = c("mouse",
                                    "human",
                                    "mouse_2DG")))

## Reorder
df <- df[order(df$time),]
```

### Get orthologous gene names for mapping between species

```{r orthologous gene names, warning = FALSE}

df_Hs2Mm <- read.csv("../gene_names_mouse_human_df.csv")

df <- rbind(df %>%
                ungroup() %>% 
                filter(Sample == "human") %>% 
                rename(gene_name_Hs = gene_name) %>% 
                left_join(df_Hs2Mm,
                          by = "gene_name_Hs"),
            df %>%
                ungroup() %>%
                filter(Sample == "mouse") %>% 
                mutate(gene_name_Mm = tolower(gene_name) %>% capitalize()) %>% 
                dplyr::select(-gene_name) %>% 
                left_join(df_Hs2Mm,
                          by = "gene_name_Mm"),
            df %>%
                ungroup() %>%
                filter(Sample == "mouse_2DG") %>% 
                mutate(gene_name_Mm = tolower(gene_name) %>% capitalize()) %>% 
                dplyr::select(-gene_name) %>% 
                left_join(df_Hs2Mm,
                          by = "gene_name_Mm")) %>% 
  ## Choose human gene name to go forward with
  mutate(gene_name = gene_name_Hs)

## For the cases, where there is no human homologue, choose the mouse gene name
df$gene_name[is.na(df$gene_name)] <- df$gene_name_Mm[is.na(df$gene_name)]

## Some special cases...
df$gene_name[grepl("^Kifc5b$", df$gene_name_Mm)] <- df$gene_name_Mm[grepl("^Kifc5b$", df$gene_name_Mm)]
df$gene_name[grepl("^Gm14325$", df$gene_name_Mm)] <- df$gene_name_Mm[grepl("^Gm14325$", df$gene_name_Mm)]

print(paste0("After mapping of gene names across species, left with ", n_distinct(df$gene_name), " gene names."))

```

### VSN normalization of Total Signal Sum (will not affect H2L ratios)

Will only be used for MA plots

```{r VSN normalization, warning = FALSE, message = FALSE}

#### ------ VSN normalization ---- ####
## Calculate normalized intensities across all samples within each experiment using vsn
list.df <- list()

## Cycle through experiments
for(i in 1:length(unique(df$Experiment))) {
  
  ## Choose only Total_Signal_Sums and convert to wide format
  df_vsn <- subset(df,
                   Experiment == unique(df$Experiment)[i],
                   select = c("gene_name", "protein_id", "source_file", "Total_Signal_Sum")) %>% 
    unique() %>% 
    arrange(gene_name)
  
  df_vsn <- spread(df_vsn, source_file, Total_Signal_Sum)
  
  
  ## Make matrix out of dataset being processed
  prot_mat <- as.matrix(df_vsn[,-1])
  
  
  ## Make all values of 0 in the signal sum to NA
  prot_mat[prot_mat == 0] <- NA

  ## Do the VSN normalization and add it back to the original data
  vsn_fit <- vsn2(prot_mat)
  # meanSdPlot(vsn_fit)
  vsn_norm <- predict(vsn_fit, prot_mat)
  
  ## Remove old values from df_vsn (only leaving gene_names) and add vsn_norm values
  df_vsn <- bind_cols(df_vsn[,1] %>% dplyr::select(everything()), as_tibble(2^vsn_norm))
  
  ## Transform back into long format
  df_vsn <- gather(df_vsn, source_file, vsn_norm_TotalSignalSum, -gene_name, na.rm = TRUE)
  
  ## Add to original data frame
  list.df[[i]] <- left_join(subset(df, Experiment == unique(df$Experiment)[i]),
                            df_vsn,
                            by = c("source_file", "gene_name"))
}

# Combine all dataframes into one
df <- do.call(rbind, list.df)

rm(prot_mat, vsn_norm, df_vsn, vsn_fit, list.df, i)

```

## Proteomic ruler quant
## Based on concept from Wisniewski et al, MCP 2014

```{r proteomic ruler quant}

## Read in histone list
histones <- read_delim("./Histone_gene_names_old_included.txt",
                       delim = "\t") %>% 
  pivot_longer(everything()) %>% 
  mutate(gene_name = strsplit(value, ",")) %>% 
  unnest(gene_name) %>% 
  na.omit() %>% 
  pull(gene_name) %>% 
  unique()

## Identify histones in data
df$Histone <- grepl(paste0(histones, collapse = "|"), df$gene_name)

## How many histones identified
df %>% 
  group_by(Sample) %>% 
  filter(Histone) %>% 
  summarise(histones = n_distinct(gene_name))

## Calculate signal sum of all histones in each source_file
df_prot_ruler <- df %>%
  left_join(
    df %>% 
      group_by(source_file) %>% 
      filter(Histone) %>% 
      dplyr::select(source_file, gene_name, vsn_norm_TotalSignalSum) %>%
      distinct() %>% 
      mutate(Histone_vsn_signal_sum = sum(vsn_norm_TotalSignalSum)) %>% 
      dplyr::select(source_file, Histone_vsn_signal_sum) %>% 
      distinct(),
    by = "source_file") %>% 
  left_join(
    df %>% 
      group_by(source_file) %>% 
      dplyr::select(source_file, gene_name, vsn_norm_TotalSignalSum) %>%
      distinct() %>%
      mutate(Run_Total_vsn_signal_sum = sum(vsn_norm_TotalSignalSum)) %>% 
      dplyr::select(source_file, Run_Total_vsn_signal_sum) %>% 
      distinct(),
    by = "source_file"
  )

## Calculate histone fractions
df_prot_ruler <- df_prot_ruler %>% 
  mutate(Hist_frac = Histone_vsn_signal_sum/Run_Total_vsn_signal_sum)

plots <- list()

## Correct for genomic size and calculate proportion of histones to other proteins per cell
## Histone content per cell

## Calculate protein copy numbers for all proteins
## First, add in mass of genomic DNA per cell
df_prot_ruler$mass_DNA <- 0
df_prot_ruler$mass_DNA[df_prot_ruler$Sample == "human"] <- 6.5389E-12 ## grams
df_prot_ruler$mass_DNA[df_prot_ruler$Sample == "mouse"] <- 5.9088E-12 ## grams
df_prot_ruler$mass_DNA[df_prot_ruler$Sample == "mouse_2DG"] <- 5.9088E-12 ## grams

## Calculate mean molecular weight per protein
df_prot_ruler <- df_prot_ruler %>%
  left_join(
    df_prot_ruler %>% 
      dplyr::select(gene_name, Sample, mw) %>% 
      distinct() %>% 
      mutate(mw = strsplit(mw, "\\|")) %>% 
      unnest(mw) %>% 
      mutate(mw = as.numeric(mw)) %>% 
      group_by(gene_name, Sample) %>% 
      mutate(mw_mean = mean(mw)) %>%
      dplyr::select(-mw) %>% 
      distinct(),
    by = c("gene_name", "Sample")
  ) %>% 
  ## Calculate sum intensity per gene name
  group_by(gene_name, Sample, time, Replicate) %>% 
  mutate(sum_vsn_norm_TotalSignalSum = sum(vsn_norm_TotalSignalSum))
  

## Calculate protein copy numbers per cell
df_prot_ruler <- df_prot_ruler %>% 
  ungroup() %>% 
  mutate(copy_number_per_cell = sum_vsn_norm_TotalSignalSum / mw_mean * mass_DNA / Histone_vsn_signal_sum * 6.022E23)
  
## Scatter plot of copy numbers per cell across species
df_plot <- df_prot_ruler %>% 
  dplyr::select(gene_name, Sample, time, Replicate, copy_number_per_cell) %>% 
  distinct()

df_plot_w <- df_plot %>% 
           pivot_wider(names_from = Sample,
                       values_from = copy_number_per_cell)

plots[[length(plots)+1]] <-
  ggplot(df_plot_w,
         aes(x = log10(human),
             y = log10(mouse))) +
    geom_density_2d() +
    theme_HH_paper +
    geom_abline() +
    ggtitle("Protein copy numbers per cell\ngene names matched") +
    facet_wrap(~time+Replicate)

plots[[length(plots)+1]] <-
  ggplot(df_plot,
         aes(x = factor(time),
             y = log10(copy_number_per_cell),
             fill = Sample)) +
    geom_boxplot() +
    theme_HH_paper +
    scale_fill_manual(values = COLOR_VECTOR_SAMPLE) +
    ggtitle("Protein copy numbers per cell") +
    facet_wrap(~Replicate)

## Median copy number per cell over replicates and timepoints
#plots[[length(plots)+1]] <-
FIG_WIDTH <- 1.5
FIG_HEIGHT <- 3
PLOT_LIMITS <- c(2, 9)

FILENAME <- "FigureS1b_Copy_numbers"

  ggplot(df_plot %>% 
           filter(Sample == "human" | Sample == "mouse") %>% 
           group_by(Sample, gene_name) %>% 
           summarise(median_copy_number_per_cell = median(copy_number_per_cell)),
         aes(x = Sample,
             y = log10(median_copy_number_per_cell),
             fill = Sample)) +
    #geom_violin() +
    geom_boxplot() +
    theme_HH_paper +
    scale_fill_manual(values = COLOR_VECTOR_SAMPLE) +
    ylim(PLOT_LIMITS) +
    stat_compare_means(
      label = "p.format",
      method = "wilcox.test",
      comparisons = list(c("human", "mouse"))) +
    ylab(bquote("log" [10] * "(Protein copy number per cell)")) +
    theme(axis.title.x = element_blank(),
          legend.position = "none") +
    stat_summary(fun.data = function(x){
  data.frame(y=min(x)*0.9,
             label = length(x))
}, geom="text")# +
    #ggtitle("Protein copy numbers per protein (gene name) per cell\nMedian over replicates and timepoints")

  ggsave(paste0(FILENAME, "_", VERSION, ".pdf"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

ggsave(paste0(FILENAME, "_", VERSION, ".svg"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)
  
write.xlsx2(df_plot %>% 
              filter(Sample == "human" | Sample == "mouse") %>% 
           group_by(Sample, gene_name) %>% 
           summarise(median_copy_number_per_cell = median(copy_number_per_cell)) %>%  
           ungroup() %>%
           as.data.frame(),
         file = paste0(FILENAME, "_", VERSION, ".xlsx"),
         sheet = "all_data",
         row.names = F)




## Look at total number of histones per cell
df_plot$Histone <- "Other proteins"
df_plot$Histone[grepl(paste0(histones, collapse = "|"), df_plot$gene_name)] <- "Histones"

df_histones <- df_plot %>% 
  group_by(Replicate, time, Sample, Histone) %>% 
  summarise(Sum_proteins = sum(copy_number_per_cell))

## Normalize to median of human protein count
df_histones <- df_histones %>% 
  mutate(human_median = df_histones %>% 
           filter(Sample == "human") %>%
           group_by(Histone) %>%
           summarise(human_median = median(Sum_proteins)) %>% 
           pull(human_median),
         logFC_human = log2(Sum_proteins/human_median))

plots[[length(plots)+1]] <-
  ggplot(df_histones,
         aes(x = Sample,
             y = logFC_human,
             fill = Sample)) +
    geom_boxplot() +
    geom_beeswarm() +
    theme_HH_paper +
    scale_fill_manual(values = COLOR_VECTOR_SAMPLE) +
    ggtitle("Sum of total number of proteins per cell,\nnormalized to human median\nEach datapoint is one MS run (timepoint, replicate)") +
    theme(axis.title.x = element_blank()) +
    facet_wrap(~Histone)
  
################

FIG_WIDTH <- 3
FIG_HEIGHT <- 3

PLOT_LIMITS <- c(6, 10.2)

# FILENAME <- "FigureS_log10_Copy_numbers_composition_human_vs_mouse"

## Median across all timepoints and replicates to match across species
df_composition_plot <- df_plot %>% 
    filter(Sample == "human" | Sample == "mouse") %>%
  ## Get sum for each sample
  group_by(Histone, Sample, Replicate, time) %>% 
  summarise(proteins_per_cell = sum(copy_number_per_cell)) %>% 
  ## Get median across replicates
  group_by(Histone, Sample) %>% 
  summarise(median_proteins_per_cell = median(proteins_per_cell),
            sd_proteins_per_cell = sd(proteins_per_cell),
            sd_log10_proteins_per_cell = sd(log10(proteins_per_cell)))


## log 10
plots[[length(plots)+1]] <-
ggplot(df_composition_plot,
       aes(x = Sample,
           y = log10(median_proteins_per_cell),
           fill = Histone)) +
  geom_col(position = position_dodge(width = 0.55),
           width = 0.5) +
  geom_errorbar(aes(ymin = log10(median_proteins_per_cell)-sd_log10_proteins_per_cell,
                    ymax = log10(median_proteins_per_cell)+sd_log10_proteins_per_cell),
                position = position_dodge(width = 0.55),
                width = 0.25) +
  coord_cartesian(ylim = PLOT_LIMITS) +
  scale_y_continuous(labels = function(x) format(10^x, scientific = TRUE),
                     breaks = c(6, 7, 8, 9, 10),
                     limits = PLOT_LIMITS,
                     oob = rescale_none) +
  theme_HH_paper +
  scale_fill_manual(values = c("Histones" = "grey20",
                               "Other proteins" = "grey")) +
  ylab(bquote("log"[10]*" proteins per cell"))

#################
## Without log scale
FIG_WIDTH <- 3
FIG_HEIGHT <- 3

PLOT_LIMITS <- c(6, 10.2)

# FILENAME <- "FigureS_Copy_numbers_composition_human_vs_mouse"

## Manually calculate errorbar positions for stacked plots)
df_errorbars <- 
  df_composition_plot %>% 
  dplyr::select(Sample, Histone, sd_proteins_per_cell) %>% 
  left_join(
  df_composition_plot %>% 
  dplyr::select(Sample, Histone, median_proteins_per_cell) %>% 
  pivot_wider(names_from = Histone,
              values_from = median_proteins_per_cell)) %>% 
  group_by(Sample) %>% 
  mutate(sd_min = 
           case_when(Histone == "Histones" ~ Histones - sd_proteins_per_cell,
                     T ~ Histones + `Other proteins` - sd_proteins_per_cell),
         sd_max = 
           case_when(Histone == "Histones" ~ Histones + sd_proteins_per_cell,
                     T ~ Histones + `Other proteins` + sd_proteins_per_cell))

plots[[length(plots)+1]] <-
ggplot(df_composition_plot %>% 
         mutate(Histone = factor(Histone, levels = c("Other proteins", "Histones"))),
       aes(x = Sample,
           fill = Histone)) +
  geom_col(aes(y = median_proteins_per_cell),
           position = "stack",
           width = 0.5) +
  geom_errorbar(data = df_errorbars,
                aes(ymin = sd_min,
                    ymax = sd_max),
                width = 0.15) +
  theme_HH_paper +
  scale_fill_manual(values = c("Histones" = "grey20",
                               "Other proteins" = "grey")) +
  ylab(bquote("sum of proteins per cell")) +
  xlab("") +
  labs(fill = "")

#################
## Gene names that can't be mapped

PLOT_LIMITS <- c(6, 10.2)

# FILENAME <- "FigureS_Copy_numbers_composition_human_vs_mouse"

## Median across all timepoints and replicates to match across species
df_composition_plot <- df_plot %>% 
    filter(Sample == "human" | Sample == "mouse") %>%
  ## Check, whether is present in both
  group_by(gene_name, Replicate, time) %>% 
  mutate(present_in_n_species = n_distinct(Sample)) %>%
  ungroup() %>% 
  mutate(present_in_species = 
           case_when(
             present_in_n_species == 2 ~ "present in both species",
             T ~ "present in only one species") %>% 
           factor(levels = c("present in only one species", "present in both species"))) %>% 
  ## Get sum for each sample
  group_by(present_in_species, Sample, Replicate, time) %>% 
  summarise(proteins_per_cell = sum(copy_number_per_cell)) %>% 
  ## Get median across replicates
  group_by(present_in_species, Sample) %>% 
  summarise(median_proteins_per_cell = median(proteins_per_cell),
            sd_proteins_per_cell = sd(proteins_per_cell),
            sd_log10_proteins_per_cell = sd(log10(proteins_per_cell)))

## Manually calculate errorbar positions for stacked plots)
df_errorbars <- 
  df_composition_plot %>% 
  dplyr::select(Sample, present_in_species, sd_proteins_per_cell) %>% 
  left_join(
  df_composition_plot %>% 
  dplyr::select(Sample, present_in_species, median_proteins_per_cell) %>% 
  pivot_wider(names_from = present_in_species,
              values_from = median_proteins_per_cell)) %>% 
  group_by(Sample) %>% 
  mutate(sd_min = 
           case_when(present_in_species == "present in both species" ~ `present in both species` - sd_proteins_per_cell,
                     T ~ `present in both species` + `present in only one species` - sd_proteins_per_cell),
         sd_max = 
           case_when(present_in_species == "present in both species" ~ `present in both species` + sd_proteins_per_cell,
                     T ~ `present in both species` + `present in only one species` + sd_proteins_per_cell))
  
plots[[length(plots)+1]] <-
ggplot(df_composition_plot,
       aes(x = Sample,
           fill = present_in_species)) +
  geom_col(aes(y = median_proteins_per_cell),
           position = "stack",
           width = 0.5) +
  geom_errorbar(data = df_errorbars,
                aes(ymin = sd_min,
                    ymax = sd_max),
                width = 0.15) +
  theme_HH_paper +
  scale_fill_manual(values = c("present in both species" = "grey20",
                               "present in only one species" = "grey")) +
  ylab(bquote("sum of proteins per cell")) +
  xlab("") +
  labs(fill = "")

ggexport(ggarrange(plotlist = plots, nrow = 1, ncol = 1, widths = 5), filename = paste0("Copy_numbers_per_cell_", VERSION, ".pdf"))
```

### Protein copy number per cell across samples

Median across timepoints and replicates

```{r}

FIG_WIDTH <- 3
FIG_HEIGHT <- 3

PLOT_LIMITS <- c(2.5, 8)

FILENAME <- "FigureS_Copy_numbers_human_vs_mouse"

## Median across all timepoints and replicates to match across species
df_median_copy_numbers <- df_plot %>% 
    filter(Sample == "human" | Sample == "mouse") %>%
  group_by(gene_name, Sample) %>% 
  summarise(log10_median_copy_number_per_cell = log10(median(copy_number_per_cell))) %>% 
  pivot_wider(names_from = Sample,
              values_from = log10_median_copy_number_per_cell) %>% 
  na.omit()

ggplot(df_median_copy_numbers,
       aes(x = human,
           y = mouse)) +
  geom_pointdensity(size = 0.5) +
  scale_color_viridis() +
  geom_abline(linewidth = 0.2) +
  coord_fixed(xlim = PLOT_LIMITS,
              ylim = PLOT_LIMITS) +
  xlab(label = bquote(~ "human, protein copy number per cell")) +
  ylab(label = bquote(~ "mouse, protein copy number per cell")) +
  theme_HH_scatter_paper +
  guides(color = FALSE) +
    geom_text(data = tibble(
    x = 6,
    y = 3),
    aes(x,y, label = paste0("N = ", n_distinct(df_median_copy_numbers$gene_name)))) +
  scale_x_continuous(labels = function(x) format(10^x, scientific = TRUE),
                     breaks = c(2, 3, 4, 5, 6, 7, 8),
                     limits = PLOT_LIMITS) +
  scale_y_continuous(labels = function(x) format(10^x, scientific = TRUE),
                     breaks = c(2, 3, 4, 5, 6, 7, 8),
                     limits = PLOT_LIMITS)

ggsave(paste0(FILENAME, "_", VERSION, ".pdf"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

ggsave(paste0(FILENAME, "_", VERSION, ".svg"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

write.xlsx2(df_median_copy_numbers %>% 
           ungroup() %>%
           as.data.frame(),
         file = paste0(FILENAME, "_", VERSION, ".xlsx"),
         sheet = "all_data",
         row.names = F)


```

# Fit turnover

Fit using a linear fit to ln(HEAVY/LIGHT+1) vs time.      
Set intercept to 0.
Fit separately for each replicate

```{r}

MIN_N_TIMEPOINTS <- 2

## Fit proteins
df_fit_proteins <- df %>%
  dplyr::select(gene_name, protein_id, Replicate, Sample, time, rel_fc_HEAVY) %>% 
  unique() %>% 
  na.omit() %>% 
  group_by(gene_name, protein_id, Replicate, Sample) %>%
  add_count() %>% 
  filter(n >= MIN_N_TIMEPOINTS) %>%
  nest() %>%
  mutate(fit = map(data, ~ lm(log(rel_fc_HEAVY+1) ~ time + 0, data = .)),
         rss = map_dbl(fit, ~deviance(.x)),
         slope = map_dbl(fit, ~coef(.x)["time"])
         ) %>% 
  ungroup() %>% 
  dplyr::select(-fit)


```

Number of gene names with fitted turnover

```{r}

print(paste0("Turnover estimates (before correction) could be fitted to ",  n_distinct(df_fit_proteins$gene_name), " gene names."))

```

## Estimate cell cycle time from data
HOX! Here, the lowest 1% of slopes is taken as a proxy for the cell cycle time

```{r, message = F, warning = F}

df_tcc_estimated <- df_fit_proteins %>% 
  group_by(Replicate, Sample) %>% 
  filter(slope > 0) %>% 
  summarise(estimated_tcc = log(2)/quantile(slope, probs = 0.01)) %>% 
  ungroup()

## Calculate half-lives
df_fit_proteins <- df_fit_proteins %>%
  ## Add tcc
  left_join(df_tcc_estimated,
            by = c("Sample", "Replicate")) %>% 
  mutate(kdeg = slope - log(2)/estimated_tcc,
         Thalf = log(2) / kdeg)

```

### Cell cycle estimate

Estimated from lowest 1% of protein turnover

```{r}

FILENAME <- "FigureS3a_CellCycleEstimate"
FIG_WIDTH <- 2.5
FIG_HEIGHT <- 3

ggplot(df_tcc_estimated,
         aes(x=Sample,
             fill = Sample)) +
  geom_col(data = . %>% 
             group_by(Sample) %>% 
             summarise(avg_tcc = mean(estimated_tcc)),
           aes(y = avg_tcc),
           alpha = 1,
           color = "black",
           linewidth = 0.3) +
  geom_errorbar(data = . %>% 
             group_by(Sample) %>% 
             summarise(avg_tcc = mean(estimated_tcc),
                       sd_tcc = sd(estimated_tcc)),
             aes(ymin = avg_tcc - sd_tcc,
                 ymax = avg_tcc + sd_tcc),
             width = 0.25) +
  geom_jitter(aes(y = estimated_tcc),
             fill = "white",
             shape = 21,
             size =3,
             width = 0.25) +
  scale_fill_manual(values = COLOR_VECTOR_SAMPLE) +
  theme_HH_paper +
  theme(legend.position = "none") +
  ylab("cell cycle estimate (h)") +
  xlab("")

ggsave(paste0(FILENAME, "_", VERSION, ".pdf"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

ggsave(paste0(FILENAME, "_", VERSION, ".svg"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

## Write out median values
write.xlsx2(df_tcc_estimated %>% 
           ungroup() %>% 
           as.data.frame(),
         file = "Figure_S1_Cell_cycle_estimates.xlsx",
         sheet = "data",
         row.names = F)

```

## Reproducibility across replicates

Collapse to gene name.

```{r}

FILENAME <- "Reproducibility_of_slopes_across_replicates"
FIG_WIDTH <- 4
FIG_HEIGHT <- 4

df_plot <- df_fit_proteins %>% 
  filter(Sample != "mouse_2DG") %>% 
  group_by(gene_name, Replicate, Sample) %>% 
  summarise(slope = median(slope)) %>% 
  ungroup()

df_plot <- df_plot %>% 
  filter(Replicate == "Rep1") %>% 
  dplyr::select(gene_name, Sample, Rep1 = slope) %>% 
  left_join(
    df_plot %>% 
      filter(Replicate %in% c("Rep2", "Rep3")) %>% 
      dplyr::select(gene_name, Sample, Replicate, Rep2Rep3 = slope)
  )

PLOT_LIMITS <- c(-7.5, 0)

ggplot(df_plot %>% 
         filter(!is.na(Replicate)),
       aes(x = log2(Rep1),
           y = log2(Rep2Rep3))) +
    geom_pointdensity(size = 0.5) +
  scale_color_viridis() +
  geom_abline(linewidth = 0.2) +
  coord_fixed(xlim = PLOT_LIMITS,
              ylim = PLOT_LIMITS) +
  xlab(label = bquote(~ "Rep1, log" ["2"] * "(k" ["deg"] * ")")) +
  ylab(label = bquote(~ "log" ["2"] * "(k" ["deg"] * ")")) +
  theme_HH_scatter_paper +
  guides(color = FALSE) +
  stat_cor(method = "pearson", 
           aes(label = ..r.label..),
           label.x = -6.5, label.y = -0.5,
           size = 3) +
  ggtitle(bquote(~ "turnover slope, uncorrected k" ["deg"])) +
  facet_grid(cols = vars(Sample),
             rows = vars(Replicate))

ggsave(paste0(FILENAME, "_", VERSION, ".pdf"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

ggsave(paste0(FILENAME, "_", VERSION, ".svg"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

## Write out median values
write.xlsx2(df_plot %>% 
           ungroup() %>% 
           as.data.frame(),
         file = paste0(FILENAME, ".xlsx"),
         sheet = "data",
         row.names = F)

```

```{r}

FILENAME <- "FigureS1b_Reproducibility_of_Thalf_across_replicates"
FIG_WIDTH <- 4
FIG_HEIGHT <- 4

df_plot <- df_fit_proteins %>% 
  filter(Sample != "mouse_2DG") %>% 
  group_by(gene_name, Replicate, Sample) %>% 
  summarise(Thalf = median(Thalf)) %>% 
  ungroup()

df_plot <- df_plot %>% 
  filter(Replicate == "Rep1") %>% 
  dplyr::select(gene_name, Sample, Rep1 = Thalf) %>% 
  left_join(
    df_plot %>% 
      filter(Replicate %in% c("Rep2", "Rep3")) %>% 
      dplyr::select(gene_name, Sample, Replicate, Rep2Rep3 = Thalf)
  )

PLOT_LIMITS <- c(0, 4)

ggplot(df_plot %>% 
         filter(!is.na(Replicate)),
       aes(x = log10(Rep1),
           y = log10(Rep2Rep3))) +
    geom_pointdensity(size = 0.5) +
  scale_color_viridis() +
  geom_abline(linewidth = 0.2) +
  coord_fixed(xlim = PLOT_LIMITS,
              ylim = PLOT_LIMITS) +
  xlab(label = bquote(~ "Rep1, log" ["10"] * "(T" ["1/2"] * ", h)")) +
  ylab(label = bquote(~ "log" ["10"] * "(T" ["1/2"] * ", h)")) +
  theme_HH_scatter_paper +
  guides(color = FALSE) +
  stat_cor(method = "pearson",
           aes(label = ..r.label..),
           label.x = 0.5, label.y = 3.5,
           size = 3) +
  ggtitle(bquote(~ "turnover, corrected T" ["1/2"] * " (h)")) +
  facet_grid(cols = vars(Sample),
             rows = vars(Replicate))

ggsave(paste0(FILENAME, "_", VERSION, ".pdf"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

ggsave(paste0(FILENAME, "_", VERSION, ".svg"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

## Write out median values
write.xlsx2(df_plot %>% 
           ungroup() %>% 
           as.data.frame(),
         file = paste0(FILENAME, ".xlsx"),
         sheet = "data",
         row.names = F)

```

## Thalf scatter plots

Use median across replicates

```{r, message = F, warning = F}

df_fit_proteins_medians <-
  df_fit_proteins %>%
  dplyr::select(gene_name, Sample, kdeg, Replicate) %>% 
  distinct() %>% 
  group_by(gene_name, Sample) %>% 
  summarise(median_kdeg = median(kdeg, na.rm = T),
            median_Thalf = log(2)/median_kdeg,
            found_in_replicates = n_distinct(Replicate)) %>% 
  ## Filter to only include cases with a minimum of two replicates for human and mouse
  filter(Sample == "mouse" & found_in_replicates > 1 |
           Sample == "human" & found_in_replicates > 1 |
           Sample == "mouse_2DG") %>% 
  mutate(Sample = factor(Sample,
                         levels = c("mouse",
                                    "human",
                                    "mouse_2DG")))

print(paste0("Turnover estimates in >=2 replicates (mouse and human) or 1 replicate (mouse DG) in total for ",  n_distinct(df_fit_proteins_medians$gene_name), " gene names."))

```

mouse v human: Only show proteins detected in at least 2 replicates.    
mouse 2DG v others: Show also cases with a single replicate

```{r, message = F, warning = F}

FIG_WIDTH <- 3
FIG_HEIGHT <- 3

PLOT_LIMITS <- c(0, 3.5)

FILENAME <- "Figure1c_Thalf_human_vs_mouse"

df_plot <- df_fit_proteins_medians %>% 
  filter(Sample == "human" | Sample == "mouse") %>% 
         dplyr::select(gene_name, Sample, median_Thalf) %>% 
  mutate(median_Thalf = log10(median_Thalf)) %>% 
         distinct() %>% 
         pivot_wider(names_from = Sample,
                     values_from = median_Thalf) %>% 
  na.omit()

ggplot(df_plot,
       aes(x = human,
           y = mouse)) +
  geom_pointdensity(size = 0.5) +
  scale_color_viridis() +
  geom_abline(linewidth = 0.2) +
  coord_fixed(xlim = PLOT_LIMITS,
              ylim = PLOT_LIMITS) +
  xlab(label = bquote(~ "human T" ["1/2"] * ", (h)")) +
  ylab(label = bquote(~ "mouse T" ["1/2"] * ", (h)")) +
  theme_HH_scatter_paper +
  guides(color = FALSE) +
    geom_text(data = tibble(
    x = 2.5,
    y = 0.5),
    aes(x,y, label = paste0("N = ", n_distinct(df_plot$gene_name)))) +
  scale_x_continuous(labels = function(x) 10^x,
                     breaks = c(0, 1, 2, 3),
                     limits = PLOT_LIMITS) +
  scale_y_continuous(labels = function(x) 10^x,
                     breaks = c(0, 1, 2, 3),
                     limits = PLOT_LIMITS)

ggsave(paste0(FILENAME, "_", VERSION, ".pdf"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

ggsave(paste0(FILENAME, "_", VERSION, ".svg"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

write.xlsx2(df_plot %>% 
           ungroup() %>%
           as.data.frame(),
         file = paste0(FILENAME, "_", VERSION, ".xlsx"),
         sheet = "all_data",
         row.names = F)


#############


FILENAME <- "Figure3f_Thalf_mouse_vs_2DG"

df_plot <- df_fit_proteins_medians %>% 
  filter(Sample == "mouse_2DG" |
           Sample == "mouse") %>% 
         dplyr::select(gene_name, Sample, median_Thalf) %>% 
  mutate(median_Thalf = log10(median_Thalf)) %>% 
         distinct() %>% 
         pivot_wider(names_from = Sample,
                     values_from = median_Thalf) %>% 
  na.omit()

ggplot(df_plot,
       aes(x = mouse_2DG,
           y = mouse)) +
  geom_pointdensity(size = 0.5) +
  scale_color_viridis() +
  geom_abline(linewidth = 0.2) +
  coord_fixed(xlim = PLOT_LIMITS,
              ylim = PLOT_LIMITS) +
  xlab(label = bquote(~ "mouse - 2DG T" ["1/2"] * ", (h)")) +
  ylab(label = bquote(~ "mouse T" ["1/2"] * ", (h)")) +
  theme_HH_scatter_paper +
  guides(color = FALSE) +
  scale_x_continuous(labels = function(x) 10^x,
                     breaks = c(0, 1, 2, 3),
                     limits = PLOT_LIMITS) +
  scale_y_continuous(labels = function(x) 10^x,
                     breaks = c(0, 1, 2, 3),
                     limits = PLOT_LIMITS) +
  geom_text(data = tibble(
    x = 2.5,
    y = 0.5),
    aes(x,y, label = paste0("N = ", n_distinct(df_plot$gene_name))))

ggsave(paste0(FILENAME, "_", VERSION, ".pdf"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

ggsave(paste0(FILENAME, "_", VERSION, ".svg"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

write.xlsx2(df_plot %>% 
           ungroup() %>%
           as.data.frame(),
         file = paste0(FILENAME, "_", VERSION, ".xlsx"),
         sheet = "all_data",
         row.names = F)

###############

FILENAME <- "Figure4a_Thalf_human_vs_2DG"

df_plot <- df_fit_proteins_medians %>% 
  filter(Sample == "mouse_2DG" |
           Sample == "human") %>% 
         dplyr::select(gene_name, Sample, median_Thalf) %>% 
  mutate(median_Thalf = log10(median_Thalf)) %>% 
         distinct() %>% 
         pivot_wider(names_from = Sample,
                     values_from = median_Thalf) %>% 
  na.omit()

ggplot(df_plot,
       aes(x = human,
           y = mouse_2DG)) +
  geom_pointdensity(size = 0.5) +
  scale_color_viridis() +
  geom_abline(linewidth = 0.2) +
  coord_fixed(xlim = PLOT_LIMITS,
              ylim = PLOT_LIMITS) +
   xlab(label = bquote(~ "human T" ["1/2"] * ", (h)")) +
  ylab(label = bquote(~ "mouse - 2DG T" ["1/2"] * ", (h)")) +
  theme_HH_scatter_paper +
  guides(color = FALSE) +
  scale_x_continuous(labels = function(x) 10^x,
                     breaks = c(0, 1, 2, 3),
                     limits = PLOT_LIMITS) +
  scale_y_continuous(labels = function(x) 10^x,
                     breaks = c(0, 1, 2, 3),
                     limits = PLOT_LIMITS) +
  geom_text(data = tibble(
    x = 2.5,
    y = 0.5),
    aes(x,y, label = paste0("N = ", n_distinct(df_plot$gene_name))))

ggsave(paste0(FILENAME, "_", VERSION, ".pdf"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

ggsave(paste0(FILENAME, "_", VERSION, ".svg"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

write.xlsx2(df_plot %>% 
           ungroup() %>%
           as.data.frame(),
         file = paste0(FILENAME, "_", VERSION, ".xlsx"),
         sheet = "all_data",
         row.names = F)

```

## Violin plots of Thalf distributions

```{r}

FIG_WIDTH <- 1.5
FIG_HEIGHT <- 3

PLOT_LIMITS <- c(-0.75, 4.1)

FILENAME <- "Figure1b_Thalf_violin_human_mouse"

ggplot(df_fit_proteins_medians %>% 
         filter(Sample == "human" | Sample == "mouse") %>% 
         dplyr::select(gene_name, Sample, median_Thalf) %>% 
         distinct(),
       aes(x = Sample,
           y = log10(median_Thalf),
           fill = Sample,
           color = Sample)) +
  theme(axis.text.x = element_text(angle = 90)) +
  geom_violin(linewidth = 0.3,
              alpha = 0.5) +
  geom_boxplot(lwd = LINEWIDTH,
               width=0.15,
               outlier.size = OUTLIER_SIZE) +
  stat_summary(geom = "crossbar", width=0.13, lwd = 0.3, fatten = 0, color="white", fun.data = function(x){ return(c(y=median(x), ymin=median(x), ymax=median(x)))}) +
  xlab("") +
  scale_y_continuous(labels = function(x) 10^x,
                     breaks = c(log10(1), log10(2), log10(5), log10(10), log10(20), log10(50), log10(100), log10(200), log10(500),  log10(1000)),
                     limits = PLOT_LIMITS) +
  ylab(bquote(~ "T" [1/2] * "(h)")) +
  stat_summary(fun.data = function(x){
    data.frame(y=max(x)+0.1,
               label = length(x))
  }, geom="text", color = "black",
  size = PLOT_TEXT_SIZE) +
  theme_HH_paper +
  scale_fill_manual(values = COLOR_VECTOR_SAMPLE) +
  scale_color_manual(values = COLOR_VECTOR_SAMPLE) +
  theme(legend.position = "none")

ggsave(paste0(FILENAME, "_", VERSION, ".pdf"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

ggsave(paste0(FILENAME, "_", VERSION, ".svg"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

## Write out median values
write.xlsx2(df_fit_proteins_medians %>% 
         filter(Sample == "human" | Sample == "mouse") %>% 
         dplyr::select(gene_name, Sample, median_kdeg_over_replicates = median_kdeg, median_Thalf_over_replicates = median_Thalf) %>% 
         distinct() %>% 
           ungroup() %>%
           mutate(log10_median_Thalf_over_replicates = log10(median_Thalf_over_replicates)) %>% 
           as.data.frame(),
         file = paste0(FILENAME, ".xlsx"),
         sheet = "all_data",
         row.names = F)

write.xlsx2(df_fit_proteins_medians %>% 
         filter(Sample == "human" | Sample == "mouse") %>% 
         dplyr::select(gene_name, Sample, median_Thalf) %>% 
         distinct() %>% 
           group_by(Sample) %>% 
           summarise(median_Thalf = median(median_Thalf),
                     log10_median_Thalf = log2(median_Thalf)) %>%
           ungroup() %>% 
           as.data.frame(),
         file = paste0(FILENAME, ".xlsx"),
         sheet = "median_Thalfs",
         row.names = F,
         append = TRUE)

###################
## Figure 4B
FILENAME <- "Figure4b_Thalf_violin_All"
FIG_WIDTH <- 1.75

ggplot(df_fit_proteins_medians %>%
         dplyr::select(gene_name, Sample, median_Thalf) %>% 
         distinct() %>% 
         mutate(Sample = factor(gsub("_", "\n", Sample),
                                levels = c("mouse", "human", "mouse\n2DG"))),
       aes(x = Sample,
           y = log10(median_Thalf),
           color = Sample,
           fill = Sample)) +
  geom_violin(linewidth = 0.3,
              alpha = 0.5) +
  geom_boxplot(lwd = LINEWIDTH,
               width=0.15,
               outlier.size = OUTLIER_SIZE) +
  stat_summary(geom = "crossbar", width=0.13, lwd = 0.3, fatten = 0, color="white", fun.data = function(x){ return(c(y=median(x), ymin=median(x), ymax=median(x)))}) +
  xlab("") +
  scale_y_continuous(labels = function(x) 10^x,
                     breaks = c(log10(1), log10(2), log10(5), log10(10), log10(20), log10(50), log10(100), log10(200), log10(500),  log10(1000)),
                     limits = PLOT_LIMITS) +
  ylab(bquote(~ "T" [1/2] * "(h)")) +
  stat_summary(fun.data = function(x){
    data.frame(y=max(x)+0.1,
               label = length(x))
  }, geom="text", color = "black",
  size = PLOT_TEXT_SIZE) +
  theme_HH_paper +
  scale_fill_manual(values = COLOR_VECTOR_SAMPLE) +
  scale_color_manual(values = COLOR_VECTOR_SAMPLE) +
  theme(legend.position = "none")

ggsave(paste0(FILENAME, "_", VERSION, ".pdf"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

ggsave(paste0(FILENAME, "_", VERSION, ".svg"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

## Write out median values
write.xlsx2(df_fit_proteins_medians %>% 
         dplyr::select(gene_name, Sample, median_kdeg_over_replicates = median_kdeg, median_Thalf_over_replicates = median_Thalf) %>% 
         distinct() %>% 
           ungroup() %>%
           mutate(log10_median_Thalf_over_replicates = log10(median_Thalf_over_replicates)) %>% 
           as.data.frame(),
         file = paste0(FILENAME, ".xlsx"),
         sheet = "all_data",
         row.names = F)

write.xlsx2(df_fit_proteins_medians %>% 
         dplyr::select(gene_name, Sample, median_Thalf) %>% 
         distinct() %>% 
           group_by(Sample) %>% 
           summarise(median_Thalf = median(median_Thalf),
                     log10_median_Thalf = log2(median_Thalf)) %>%
           ungroup() %>% 
           as.data.frame(),
         file = paste0(FILENAME, ".xlsx"),
         sheet = "median_Thalfs",
         row.names = F,
         append = TRUE)
```

# Compare turnover using competing fits and F-statistic
## human vs mouse

For finding differentially turning over proteins.    
Adjust for cell cycle

```{r}

## Adjust for cell cycle
df <- df %>% 
  left_join(df_tcc_estimated,
            by = c("Sample", "Replicate")) %>% 
  mutate(new2old_tcc_adjusted = log(rel_fc_HEAVY + 1) - log(2)/estimated_tcc*time)

```

human vs mouse

```{r, message = F, warning = F}

## Threshhold for, how many datapoints should be available per sample
MIN_N_DATAPOINTS <- 3
MIN_TOTAL_DATAPOINTS <- 8
## Fix intercept?
FIX_INTERCEPT = TRUE
FIXED_INTERCEPT = 0

ADJ_P_CUTOFF <- 0.001

## bin based on number of measured datapoints
NUMBER_OF_BINS <- 4

set.seed(123)

df_temp <- df %>% 
  dplyr::select(gene_name, time, Replicate, Sample, new2old_tcc_adjusted) %>% 
  unique() %>% 
  filter(Sample == "human" |
           Sample == "mouse") %>% 
  ## Check, how many datapoints available and filter
  group_by(gene_name, Sample) %>% 
  add_count(name = "datapoints_per_sample") %>%
  filter(datapoints_per_sample >= MIN_N_DATAPOINTS) %>%
  ## Then filter for cases, which have been found in more than 1 sample
  group_by(gene_name) %>%
  nest() %>% 
  mutate(detected_in_samples = map_int(data,
                             ~ .x %>% pull(Sample) %>% n_distinct)) %>%
  unnest(data) %>% 
  filter(detected_in_samples > 1) %>%
  ## Count number of datapoints in total
  group_by(gene_name) %>% 
  add_count(name = "datapoints_in_total") %>%
  filter(datapoints_in_total >= MIN_TOTAL_DATAPOINTS) %>% 
  ## Shuffle species assignment to datapoints per protein
  mutate(Sample_randomized = sample(Sample)) %>% 
  ungroup()

df_temp <- df_temp[order(df_temp$gene_name),]


## Define function to fit competing models
f_fit_competing_linear_models <- 
  function(intercept_fixed = FIX_INTERCEPT,
           fixed_intercept = FIXED_INTERCEPT,
           fdf,
           grouping_variable = "^Sample$") {
    
    if(intercept_fixed == TRUE) { ## If intercepts are to be fixed
      ## Define grouping variable
      colnames(fdf)[grepl(grouping_variable, colnames(fdf))] <- "grouping_variable"
      
      ## H0, both samples together
      H0_output <-
        fdf %>% 
          mutate(fixed_intercept = fixed_intercept) %>%
          group_by(gene_name) %>%
          nest() %>% 
          mutate(H0_model = map(data,
                             ~ lm(new2old_tcc_adjusted ~ time + 0, data = .x, offset = fixed_intercept)),
                 H0_r2 = map_dbl(H0_model,
                          ~ summary(.x)$r.squared),
                 H0_slope = map_dbl(H0_model,
                                    ~coef(.x)),
                 H0_rss = map_dbl(H0_model,
                                  ~ deviance(.x)),
                 H0_intercept = fixed_intercept,
                 H0_df.residual = map_int(H0_model,
                                         ~summary.lm(.x)$df[2])) ## These residuals are actually not used, but still extracted for reference
      
      ## H1, samples separately
      H1_output <-
        fdf %>% 
          mutate(fixed_intercept = fixed_intercept) %>%
          group_by(gene_name, grouping_variable) %>%
          nest() %>% 
          mutate(H1_model = map(data,
                             ~ lm(new2old_tcc_adjusted ~ time + 0, data = .x, offset = fixed_intercept)),
                 H1_r2 = map_dbl(H1_model,
                          ~ summary(.x)$r.squared),
                 H1_slope = map_dbl(H1_model,
                                    ~coef(.x)),
                 H1_rss = map_dbl(H1_model,
                                  ~ deviance(.x)),
                 H1_intercept = fixed_intercept,
                 H1_df.residual = map_int(H1_model,
                                          ~summary.lm(.x)$df[2]))
      
    } else {
      H0_output <-
        fdf %>%
          group_by(gene_name) %>%
          nest() %>% 
          mutate(H0_model = map(data,
                             ~ lm(new2old_tcc_adjusted ~ time, data = .x)),
                 H0_r2 = map_dbl(H0_model,
                          ~ summary(.x)$r.squared),
                 H0_slope = map_dbl(H0_model,
                                    ~coef(.x)["time"]),
                 H0_rss = map_dbl(H0_model,
                                  ~ deviance(.x)),
                 H0_intercept = map_dbl(H0_model,
                                    ~coef(.x)["(Intercept)"]),
                 H0_df.residual = map_int(H0_model,
                                         ~summary.lm(.x)$df[2]))
      
      ## H1, fitting separately
      H1_output <-
        fdf %>%
          group_by(gene_name, grouping_variable) %>%
          nest() %>% 
          mutate(H1_model = map(data,
                             ~ lm(new2old_tcc_adjusted ~ time, data = .x)),
                 H1_r2 = map_dbl(H1_model,
                          ~ summary(.x)$r.squared),
                 H1_slope = map_dbl(H1_model,
                                    ~coef(.x)["time"]),
                 H1_rss = map_dbl(H1_model,
                                  ~ deviance(.x)),
                 H1_intercept = map_dbl(H1_model,
                                    ~coef(.x)["(Intercept)"]),
                 H1_df.residual = map_int(H1_model,
                                          ~summary.lm(.x)$df[2]))
    }
  
  ## Extract and combine outputs
  fdf_combined_out <-
    left_join(H1_output %>%
                dplyr::select(gene_name, H1_rss, H1_slope, H1_intercept, H1_df.residual, grouping_variable) %>% 
                group_by(gene_name) %>% 
                mutate(H1_rss_sum = sum(H1_rss),
                       H1_df.residual_sum = sum(H1_df.residual),
                       Thalf_H1 = log(2)/H1_slope),
              H0_output %>%
                dplyr::select(gene_name, H0_rss, H0_slope, H0_intercept, H0_df.residual),
              by = "gene_name")
  
  return(fdf_combined_out)
}


model_output_randomized <-
  f_fit_competing_linear_models(intercept_fixed = FIX_INTERCEPT,
                                fixed_intercept = FIXED_INTERCEPT,
                                fdf = df_temp,
                                grouping_variable = "^Sample_randomized$")

plots <- list()

## RSS from H0 and H1 from randomized data should be practically identical
plots[[length(plots)+1]] <-
  ggplot(model_output_randomized %>% 
           dplyr::select(gene_name, H0_rss, H1_rss_sum) %>% 
           unique(),
         aes(x = H0_rss,
             y = H1_rss_sum)) +
    geom_point() +
    geom_abline(linetype = "dashed") +
    theme_classic() +
    theme(panel.background = element_rect(fill = "transparent")) +
    ggtitle("Samples randomized")

## Also slopes should be practically identical
plots[[length(plots)+1]] <-
  ggplot(model_output_randomized,
         aes(x = H0_slope,
             y = H1_slope)) +
    geom_point() +
    geom_abline(linetype = "dashed") +
    theme_classic() +
    theme(panel.background = element_rect(fill = "transparent")) +
    ggtitle("Samples randomized")

## Make function to estimate dof's
f_estimate_dof <- function(fdf) {
  ## Will follow the logic outlined in Childs et al, Mol Cell Prot 2019
  RSS_0 <- fdf$H0_rss
  RSS_1 <- fdf$H1_rss_sum
  
  ## Estimate scale paremeter sigma^2
  s2 <- 1/2 * mad(RSS_0 - RSS_1)^2 / median(RSS_0 - RSS_1)
  
  ## Find d1 and d2
  RSS_diff_per_s2 <- (RSS_0 - RSS_1) / s2
  
  ## Remove cases, which are negative or zero
  RSS_diff_per_s2 <- RSS_diff_per_s2[!(is.na(RSS_diff_per_s2) | RSS_diff_per_s2 <= 0)]
  
  ## Fit to chi-square distribution
  d1 <- fitdistr(x = RSS_diff_per_s2,
                 densfun = "chi-squared",
                 start = list(df = 1),
                 method = "Brent",
                 lower = 0,
                 upper = length(RSS_diff_per_s2))
  
  ## Extract estimate for degree of freedom
  d1 <- d1[["estimate"]]
  
  
  ## And for the second degree of freedom
  RSS1_per_s2 <- RSS_1 / s2
  
  ## Remove cases, which are negative or zero
  RSS1_per_s2 <- RSS1_per_s2[!(is.na(RSS1_per_s2) | RSS1_per_s2 <= 0)]
  
  ## Fit to chi-square distribution
  d2 <- fitdistr(x = RSS1_per_s2,
                 densfun = "chi-squared",
                 start = list(df = 1),
                 method = "Brent",
                 lower = 0,
                 upper = length(RSS1_per_s2))
  
  ## Extract estimate for degree of freedom
  d2 <- d2[["estimate"]]
  
  return(data.frame(d1 = d1,
                    d2 = d2))
}



## Bin data based on N_datapoints to estimate separate degrees of freedom for each bin
df_bins <- df_temp %>%
  dplyr::select(gene_name, datapoints_per_sample, Sample_randomized) %>% 
  unique() %>% 
  group_by(gene_name) %>% 
  summarise(mean_n_datapoints_per_sample = mean(datapoints_per_sample)) %>% 
  mutate(bin = cut(mean_n_datapoints_per_sample, NUMBER_OF_BINS) %>% as.numeric()) %>% 
  ungroup()

dofs_from_randomized <- model_output_randomized %>%
  dplyr::select(gene_name, H0_rss, H1_rss_sum, H0_slope) %>% 
  unique() %>% 
  left_join(df_bins,
            by = "gene_name")

## In randomized data difference should follow a random distribution. In practice, a chi-square distribution of some order (since adding differing numbers of RSS values to each other?)
plots[[length(plots)+1]] <-
  ggplot(dofs_from_randomized %>% 
           dplyr::select(gene_name, H0_rss, H1_rss_sum, bin) %>% 
           unique(),
         aes(x = H0_rss - H1_rss_sum,
             y = after_stat(count))) +
    geom_histogram(bins = 200) +
    theme_classic() +
    theme(panel.background = element_rect(fill = "transparent")) +
    ggtitle("Samples randomized, distribution for d1") +
    facet_wrap(~bin, nrow = 1)

plots[[length(plots)+1]] <-
  ggplot(dofs_from_randomized %>% 
           dplyr::select(gene_name, H0_rss, H1_rss_sum, bin) %>% 
           unique(),
         aes(x = H1_rss_sum,
             y = after_stat(count))) +
    geom_histogram(bins = 200) +
    theme_classic() +
    theme(panel.background = element_rect(fill = "transparent")) +
    ggtitle("Samples randomized, distribution for d2") +
    facet_wrap(~bin, nrow = 1)

#### Estimating degrees of freedom ####
## To calculate the F-statistic, the degrees of freedom need to be estimated, due to heteroscedasticity of the data
plots[[length(plots)+1]] <-
  ggplot(dofs_from_randomized %>% 
           dplyr::select(gene_name, H0_rss, H1_rss_sum, H0_slope, bin) %>% 
           unique(),
         aes(x = H0_slope,
             y = H0_rss-H1_rss_sum)) +
    geom_point(alpha = 0.3) +
    theme_classic() +
    theme(panel.background = element_rect(fill = "transparent")) +
    ggtitle("Samples randomized") +
    facet_wrap(~bin, nrow = 1)

## Find dof-estimates for the bins
dofs_from_randomized <- dofs_from_randomized %>%
  dplyr::select(gene_name, H0_rss, H1_rss_sum, bin) %>% 
  unique() %>% 
  group_by(bin) %>% 
  nest() %>% 
  mutate(dof = map(.x = data,
                   .f = f_estimate_dof)) %>%
  unnest(cols = c(dof, data)) %>% 
  ungroup() %>% 
  dplyr::select(gene_name, d1, d2, bin)

## With these, go for the F-statistic calculation in the un-randomized, proper data
## Fit competing models
model_output_proper <- 
  f_fit_competing_linear_models(intercept_fixed = FIX_INTERCEPT,
                                fixed_intercept = FIXED_INTERCEPT,
                                fdf = df_temp,
                                grouping_variable = "^Sample$")

## Add in degrees of freedom estimated from the randomized dataset
df_F_stats_out <- model_output_proper %>%
  dplyr::select(gene_name, H0_rss, H1_rss_sum, H0_df.residual, H1_df.residual_sum) %>% 
  unique() %>% 
  ungroup() %>% 
  left_join(dofs_from_randomized,
            by = "gene_name") %>% 
    ## Calculate F-statistic
  mutate(F_statistic = d2/d1 * (H0_rss - H1_rss_sum) / H1_rss_sum) %>% 
    ## Calculate P_value from F-statistic using the estimated degrees of freedom
  mutate(P_value = 1 - pf(F_statistic,
                    df1 = d1,
                    df2 = d2),
          ## Adjust P_value for multiple testing
         Adj_P_value = p.adjust(P_value,
                          method = "BH")) %>%
    ## Add in info on number of measured datapoints
  left_join(df_bins %>% 
              dplyr::select(gene_name, mean_n_datapoints_per_sample),
            by = "gene_name")



#### ---- Plots for checking, etc. ---  ####
plots[[length(plots)+1]] <-
  ggplot(df_F_stats_out,
         aes(x = F_statistic,
             y = after_stat(count))) +
    theme_classic() +
    theme(panel.background = element_rect(fill = "transparent")) +
    geom_histogram(bins = 1000)

plots[[length(plots)+1]] <-
  ggplot(df_F_stats_out,
         aes(x = log10(F_statistic),
             y = after_stat(count))) +
    theme_classic() +
    theme(panel.background = element_rect(fill = "transparent")) +
    geom_histogram(bins = 200)

plots[[length(plots)+1]] <-
  ggplot(df_F_stats_out,
         aes(x = P_value,
             y = after_stat(count))) +
    theme_classic() +
    theme(panel.background = element_rect(fill = "transparent")) +
    geom_histogram(bins = 1000)

plots[[length(plots)+1]] <-
  ggplot(df_F_stats_out,
         aes(x = P_value,
             y = mean_n_datapoints_per_sample)) +
    geom_point() +
    ggtitle("Mean number of datapoints per sample") +
    theme_classic() +
    theme(panel.background = element_rect(fill = "transparent"))

plots[[length(plots)+1]] <-
  ggplot(df_F_stats_out,
         aes(x = log10(H0_rss - H1_rss_sum),
             y = -log10(P_value),
             color = factor(bin))) +
    geom_point() +
    theme_classic() +
    theme(panel.background = element_rect(fill = "transparent")) +
    ggtitle("Bins with number of datapoints per sample")

plots[[length(plots)+1]] <-
  ggplot(df_F_stats_out,
         aes(x = log10(P_value),
             y = log10(Adj_P_value),
             color = factor(bin))) +
    geom_point() +
    theme_classic() +
    theme(panel.background = element_rect(fill = "transparent")) +
    geom_abline(linetype = "dashed")

## Example fits for representative purposes
plots[[length(plots)+1]] <-
  ggplot(df_temp %>% 
           filter(gene_name == "ARF4" | gene_name == "PSMD13" | gene_name == "ALDOA"),
         aes(x = time,
             y = new2old_tcc_adjusted,
             color = Sample)) +
    geom_point(size = 3) +
    geom_smooth(method = "lm",
                formula = y ~ x+0,
                se = FALSE) +
    theme_classic() +
    theme(panel.background = element_rect(fill = "transparent"),
          axis.text=element_text(size=10)) +
    ggtitle("H1") +
    scale_color_manual(values = COLOR_VECTOR_SAMPLE) +
    facet_wrap(~gene_name) +
  theme(panel.spacing = unit(2, "lines"))

plots[[length(plots)+1]] <-
  ggplot(df_temp %>% 
           filter(gene_name == "ARF4" | gene_name == "PSMD13" | gene_name == "ALDOA"),
         aes(x = time,
             y = new2old_tcc_adjusted)) +
    geom_point(size = 3) +
    geom_smooth(method = "lm",
                formula = y ~ x+0,
                se = FALSE) +
    theme_classic() +
    theme(panel.background = element_rect(fill = "transparent"),
          axis.text=element_text(size=10)) +
    ggtitle("H0") +
    facet_wrap(~gene_name) +
  theme(panel.spacing = unit(2, "lines"))

ggexport(ggarrange(plotlist = plots, nrow = 1, ncol = 1, widths = 12),
         filename = paste0("F-statistics_competing_fits_human_v_mouse_", VERSION, ".pdf"),
         height = 4, width = 8)

rm(df_bins, df_temp, plots, dofs_from_randomized, model_output_proper, model_output_randomized)
```


```{r}
## Add in information, whether slower or faster
df_F_stats_human_v_mouse <-
  df_fit_proteins_medians %>% 
  dplyr::select(-c(found_in_replicates, median_Thalf)) %>% 
  ## Remove negative kdegs
  filter(median_kdeg > 0) %>% 
  pivot_wider(names_from = Sample,
              values_from = median_kdeg) %>% 
  dplyr::select(gene_name, mouse, human) %>% 
  mutate(human_vs_mouse = human - mouse) %>% 
  left_join(df_F_stats_out %>% 
              dplyr::select(gene_name, H0_rss, H1_rss_sum, F_statistic, Adj_P_value)) %>% 
  filter(!is.na(F_statistic) &
           !is.na(human) &
           !is.na(mouse)) %>% 
  ungroup() %>% 
  mutate(Hit_type = case_when(
    Adj_P_value <= ADJ_P_CUTOFF & human_vs_mouse > 0 ~ "faster in human",
    Adj_P_value <= ADJ_P_CUTOFF & human_vs_mouse < 0 ~ "slower in human",
    T ~ "no difference"), 
    Hit_type = factor(Hit_type,
                      levels = c("no difference",
                                 "faster in human",
                                 "slower in human")))
#rm(df_F_stats_out)

## How many in each category?
df_F_stats_human_v_mouse %>% 
  dplyr::select(Hit_type) %>% 
  table()

```

## Volcano-like plot from F statistics: human v mouse

```{r, message = F, warning = F}
FIG_WIDTH <- 3
FIG_HEIGHT <- 4

# PLOT_LIMITS <- c(-14, 0)

FILENAME <- "Figure1d_Volcano_human_mouse"

ggplot(df_F_stats_human_v_mouse,
       aes(x = sign(human_vs_mouse)*sqrt(H0_rss - H1_rss_sum),
           y = log2(F_statistic + 1))) +
  geom_point(shape = 1,
             size = 0.8,
             stroke = 0.4,
             aes(color = Hit_type)) + 
  scale_color_manual(values = COLOR_HITS) +
  ylab(bquote(~ "log"["2"]*"(" * italic("F") *  "-statistic + 1)")) +
  # xlab(bquote(~ "sign(" * Delta * "initial slope)" * sqrt("RSS"[0]*"-"*"RSS"[1]))) +
  xlab("effect size") +
  labs(color = "Hit") +
  theme_HH_scatter_paper +
  geom_text(data = tibble(
    x = 0.5,
    y = 1),
    aes(x,y, label = paste0("N = ", n_distinct(df_F_stats_human_v_mouse$gene_name)))) + 
  theme(
    legend.position = "bottom",
    legend.direction = "vertical"
  )

ggsave(paste0(FILENAME, "_", VERSION, ".pdf"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

ggsave(paste0(FILENAME, "_", VERSION, ".svg"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

write.xlsx2(df_F_stats_human_v_mouse %>%
              mutate(effect_size = sign(human_vs_mouse)*sqrt(H0_rss - H1_rss_sum),
                     y = log2(F_statistic + 1)) %>% 
           as.data.frame(),
         file = paste0(FILENAME, ".xlsx"),
         sheet = "Volcano_like",
         row.names = F)


```

## F statistic and abundance - human v mouse

```{r, message = F, warning = F}

df_plot <- df_F_stats_human_v_mouse %>% 
         mutate(log2_human_vs_mouse_Thalf = log2((log(2)/human)/(log(2)/mouse))) %>% 
  na.omit() %>% 
  left_join(
    df %>% 
      filter(Sample != "mouse_2DG") %>%
      group_by(gene_name) %>% 
      summarise(median_abundance = log10(median(vsn_norm_TotalSignalSum, na.rm = T)))
  ) %>% 
  dplyr::select(gene_name, log2_human_vs_mouse_Thalf, Hit_type, median_abundance)

FIG_WIDTH <- 3
FIG_HEIGHT <- 3

FILENAME <- "Figure1f_MA_plot_human_v_mouse"

ggplot(df_plot,
       aes(x = median_abundance,
           y = log2_human_vs_mouse_Thalf)) +
  geom_pointdensity(size = 0.5) +
  geom_hline(yintercept = 0,
             linewidth = 0.2) +
  scale_color_viridis() +
  ylab(label = bquote(~ "log" ["2"] * "(half life human / mouse)")) +
  xlab(label = bquote(~ "log" ["10"] * "(median protein abundance, a.u.)")) +
  theme_HH_scatter_paper +
  geom_text(data = tibble(
    x = 8,
    y = -4),
    aes(x,y, label = paste0("N = ", n_distinct(df_plot$gene_name)))) + 
  guides(color = FALSE)

ggsave(paste0(FILENAME, "_", VERSION, ".pdf"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

ggsave(paste0(FILENAME, "_", VERSION, ".svg"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

write.xlsx2(df_plot %>% 
           as.data.frame(),
         file = paste0(FILENAME, ".xlsx"),
         sheet = "MA_plot",
         row.names = F)



## Highlight hits from F statistic
FILENAME <- "FigureS3b_MA_plot_human_v_mouse_with_Hits"

FIG_WIDTH <- 5
FIG_HEIGHT <- 3

ggplot(df_plot %>% 
         arrange(Hit_type),
       aes(x = median_abundance,
           y = log2_human_vs_mouse_Thalf)) +
  geom_point(size = 0.5,
             aes(color = Hit_type)) +
  geom_hline(yintercept = 0,
             linewidth = 0.2) +
  scale_color_manual(values = COLOR_HITS) +
  ylab(label = bquote(~ "log" ["2"] * "(half life human / mouse)")) +
  xlab(label = bquote(~ "log" ["10"] * "(median protein abundance, a.u.)")) +
  theme_HH_scatter_paper +
  geom_text(data = tibble(
    x = 8,
    y = -4),
    aes(x,y, label = paste0("N = ", n_distinct(df_plot$gene_name)))) +
  labs(color = "Hit")

ggsave(paste0(FILENAME, "_", VERSION, ".pdf"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

ggsave(paste0(FILENAME, "_", VERSION, ".svg"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

write.xlsx2(df_plot %>% 
           as.data.frame(),
         file = paste0(FILENAME, ".xlsx"),
         sheet = "MA_plot",
         row.names = F)

```

## mouse 2DG vs mouse

For finding differentially turning over proteins

```{r, message = F, warning = F}

## Threshhold for, how many datapoints should be available per sample
MIN_N_DATAPOINTS <- 3
MIN_TOTAL_DATAPOINTS <- 6
## Fix intercept?
FIX_INTERCEPT = TRUE
FIXED_INTERCEPT = 0

ADJ_P_CUTOFF <- 0.001

## bin based on number of measured datapoints
NUMBER_OF_BINS <- 4

set.seed(123)

df_temp <- df %>% 
  dplyr::select(gene_name, time, Replicate, Sample, new2old_tcc_adjusted) %>% 
  unique() %>% 
  filter(Sample == "mouse_2DG" |
           Sample == "mouse") %>% 
  ## Check, how many datapoints available and filter
  group_by(gene_name, Sample) %>% 
  add_count(name = "datapoints_per_sample") %>%
  filter(datapoints_per_sample >= MIN_N_DATAPOINTS) %>%
  ## Then filter for cases, which have been found in more than 1 sample
  group_by(gene_name) %>%
  nest() %>% 
  mutate(detected_in_samples = map_int(data,
                             ~ .x %>% pull(Sample) %>% n_distinct)) %>%
  unnest(data) %>% 
  filter(detected_in_samples > 1) %>%
  ## Count number of datapoints in total
  group_by(gene_name) %>% 
  add_count(name = "datapoints_in_total") %>%
  filter(datapoints_in_total >= MIN_TOTAL_DATAPOINTS) %>% 
  ## Shuffle species assignment to datapoints per protein
  mutate(Sample_randomized = sample(Sample)) %>% 
  ungroup()

df_temp <- df_temp[order(df_temp$gene_name),]


## Define function to fit competing models
f_fit_competing_linear_models <- 
  function(intercept_fixed = FIX_INTERCEPT,
           fixed_intercept = FIXED_INTERCEPT,
           fdf,
           grouping_variable = "^Sample$") {
    
    if(intercept_fixed == TRUE) { ## If intercepts are to be fixed
      ## Define grouping variable
      colnames(fdf)[grepl(grouping_variable, colnames(fdf))] <- "grouping_variable"
      
      ## H0, both samples together
      H0_output <-
        fdf %>% 
          mutate(fixed_intercept = fixed_intercept) %>%
          group_by(gene_name) %>%
          nest() %>% 
          mutate(H0_model = map(data,
                             ~ lm(new2old_tcc_adjusted ~ time + 0, data = .x, offset = fixed_intercept)),
                 H0_r2 = map_dbl(H0_model,
                          ~ summary(.x)$r.squared),
                 H0_slope = map_dbl(H0_model,
                                    ~coef(.x)),
                 H0_rss = map_dbl(H0_model,
                                  ~ deviance(.x)),
                 H0_intercept = fixed_intercept,
                 H0_df.residual = map_int(H0_model,
                                         ~summary.lm(.x)$df[2])) ## These residuals are actually not used, but still extracted for reference
      
      ## H1, samples separately
      H1_output <-
        fdf %>% 
          mutate(fixed_intercept = fixed_intercept) %>%
          group_by(gene_name, grouping_variable) %>%
          nest() %>% 
          mutate(H1_model = map(data,
                             ~ lm(new2old_tcc_adjusted ~ time + 0, data = .x, offset = fixed_intercept)),
                 H1_r2 = map_dbl(H1_model,
                          ~ summary(.x)$r.squared),
                 H1_slope = map_dbl(H1_model,
                                    ~coef(.x)),
                 H1_rss = map_dbl(H1_model,
                                  ~ deviance(.x)),
                 H1_intercept = fixed_intercept,
                 H1_df.residual = map_int(H1_model,
                                          ~summary.lm(.x)$df[2]))
      
    } else {
      H0_output <-
        fdf %>%
          group_by(gene_name) %>%
          nest() %>% 
          mutate(H0_model = map(data,
                             ~ lm(new2old_tcc_adjusted ~ time, data = .x)),
                 H0_r2 = map_dbl(H0_model,
                          ~ summary(.x)$r.squared),
                 H0_slope = map_dbl(H0_model,
                                    ~coef(.x)["time"]),
                 H0_rss = map_dbl(H0_model,
                                  ~ deviance(.x)),
                 H0_intercept = map_dbl(H0_model,
                                    ~coef(.x)["(Intercept)"]),
                 H0_df.residual = map_int(H0_model,
                                         ~summary.lm(.x)$df[2]))
      
      ## H1, fitting separately
      H1_output <-
        fdf %>%
          group_by(gene_name, grouping_variable) %>%
          nest() %>% 
          mutate(H1_model = map(data,
                             ~ lm(new2old_tcc_adjusted ~ time, data = .x)),
                 H1_r2 = map_dbl(H1_model,
                          ~ summary(.x)$r.squared),
                 H1_slope = map_dbl(H1_model,
                                    ~coef(.x)["time"]),
                 H1_rss = map_dbl(H1_model,
                                  ~ deviance(.x)),
                 H1_intercept = map_dbl(H1_model,
                                    ~coef(.x)["(Intercept)"]),
                 H1_df.residual = map_int(H1_model,
                                          ~summary.lm(.x)$df[2]))
    }
  
  ## Extract and combine outputs
  fdf_combined_out <-
    left_join(H1_output %>%
                dplyr::select(gene_name, H1_rss, H1_slope, H1_intercept, H1_df.residual, grouping_variable) %>% 
                group_by(gene_name) %>% 
                mutate(H1_rss_sum = sum(H1_rss),
                       H1_df.residual_sum = sum(H1_df.residual),
                       Thalf_H1 = log(2)/H1_slope),
              H0_output %>%
                dplyr::select(gene_name, H0_rss, H0_slope, H0_intercept, H0_df.residual),
              by = "gene_name")
  
  return(fdf_combined_out)
}


model_output_randomized <-
  f_fit_competing_linear_models(intercept_fixed = FIX_INTERCEPT,
                                fixed_intercept = FIXED_INTERCEPT,
                                fdf = df_temp,
                                grouping_variable = "^Sample_randomized$")

plots <- list()

## RSS from H0 and H1 from randomized data should be practically identical
plots[[length(plots)+1]] <-
  ggplot(model_output_randomized %>% 
           dplyr::select(gene_name, H0_rss, H1_rss_sum) %>% 
           unique(),
         aes(x = H0_rss,
             y = H1_rss_sum)) +
    geom_point() +
    geom_abline(linetype = "dashed") +
    theme_classic() +
    theme(panel.background = element_rect(fill = "transparent")) +
    ggtitle("Samples randomized")

## Also slopes should be practically identical
plots[[length(plots)+1]] <-
  ggplot(model_output_randomized,
         aes(x = H0_slope,
             y = H1_slope)) +
    geom_point() +
    geom_abline(linetype = "dashed") +
    theme_classic() +
    theme(panel.background = element_rect(fill = "transparent")) +
    ggtitle("Samples randomized")

## Make function to estimate dof's
f_estimate_dof <- function(fdf) {
  ## Will follow the logic outlined in Childs et al, Mol Cell Prot 2019
  RSS_0 <- fdf$H0_rss
  RSS_1 <- fdf$H1_rss_sum
  
  ## Estimate scale paremeter sigma^2
  s2 <- 1/2 * mad(RSS_0 - RSS_1)^2 / median(RSS_0 - RSS_1)
  
  ## Find d1 and d2
  RSS_diff_per_s2 <- (RSS_0 - RSS_1) / s2
  
  ## Remove cases, which are negative or zero
  RSS_diff_per_s2 <- RSS_diff_per_s2[!(is.na(RSS_diff_per_s2) | RSS_diff_per_s2 <= 0)]
  
  ## Fit to chi-square distribution
  d1 <- fitdistr(x = RSS_diff_per_s2,
                 densfun = "chi-squared",
                 start = list(df = 1),
                 method = "Brent",
                 lower = 0,
                 upper = length(RSS_diff_per_s2))
  
  ## Extract estimate for degree of freedom
  d1 <- d1[["estimate"]]
  
  
  ## And for the second degree of freedom
  RSS1_per_s2 <- RSS_1 / s2
  
  ## Remove cases, which are negative or zero
  RSS1_per_s2 <- RSS1_per_s2[!(is.na(RSS1_per_s2) | RSS1_per_s2 <= 0)]
  
  ## Fit to chi-square distribution
  d2 <- fitdistr(x = RSS1_per_s2,
                 densfun = "chi-squared",
                 start = list(df = 1),
                 method = "Brent",
                 lower = 0,
                 upper = length(RSS1_per_s2))
  
  ## Extract estimate for degree of freedom
  d2 <- d2[["estimate"]]
  
  return(data.frame(d1 = d1,
                    d2 = d2))
}



## Bin data based on N_datapoints to estimate separate degrees of freedom for each bin
df_bins <- df_temp %>%
  dplyr::select(gene_name, datapoints_per_sample, Sample_randomized) %>% 
  unique() %>% 
  group_by(gene_name) %>% 
  summarise(mean_n_datapoints_per_sample = mean(datapoints_per_sample)) %>% 
  mutate(bin = cut(mean_n_datapoints_per_sample, NUMBER_OF_BINS) %>% as.numeric()) %>% 
  ungroup()

dofs_from_randomized <- model_output_randomized %>%
  dplyr::select(gene_name, H0_rss, H1_rss_sum, H0_slope) %>% 
  unique() %>% 
  left_join(df_bins,
            by = "gene_name")

## In randomized data difference should follow a random distribution. In practice, a chi-square distribution of some order (since adding differing numbers of RSS values to each other?)
plots[[length(plots)+1]] <-
  ggplot(dofs_from_randomized %>% 
           dplyr::select(gene_name, H0_rss, H1_rss_sum, bin) %>% 
           unique(),
         aes(x = H0_rss - H1_rss_sum,
             y = after_stat(count))) +
    geom_histogram(bins = 200) +
    theme_classic() +
    theme(panel.background = element_rect(fill = "transparent")) +
    ggtitle("Samples randomized, distribution for d1") +
    facet_wrap(~bin, nrow = 1)

plots[[length(plots)+1]] <-
  ggplot(dofs_from_randomized %>% 
           dplyr::select(gene_name, H0_rss, H1_rss_sum, bin) %>% 
           unique(),
         aes(x = H1_rss_sum,
             y = after_stat(count))) +
    geom_histogram(bins = 200) +
    theme_classic() +
    theme(panel.background = element_rect(fill = "transparent")) +
    ggtitle("Samples randomized, distribution for d2") +
    facet_wrap(~bin, nrow = 1)

#### Estimating degrees of freedom ####
## To calculate the F-statistic, the degrees of freedom need to be estimated, due to heteroscedasticity of the data
plots[[length(plots)+1]] <-
  ggplot(dofs_from_randomized %>% 
           dplyr::select(gene_name, H0_rss, H1_rss_sum, H0_slope, bin) %>% 
           unique(),
         aes(x = H0_slope,
             y = H0_rss-H1_rss_sum)) +
    geom_point(alpha = 0.3) +
    theme_classic() +
    theme(panel.background = element_rect(fill = "transparent")) +
    ggtitle("Samples randomized") +
    facet_wrap(~bin, nrow = 1)

## Find dof-estimates for the bins
dofs_from_randomized <- dofs_from_randomized %>%
  dplyr::select(gene_name, H0_rss, H1_rss_sum, bin) %>% 
  unique() %>% 
  group_by(bin) %>% 
  nest() %>% 
  mutate(dof = map(.x = data,
                   .f = f_estimate_dof)) %>%
  unnest(cols = c(dof, data)) %>% 
  ungroup() %>% 
  dplyr::select(gene_name, d1, d2, bin)

## With these, go for the F-statistic calculation in the un-randomized, proper data
## Fit competing models
model_output_proper <- 
  f_fit_competing_linear_models(intercept_fixed = FIX_INTERCEPT,
                                fixed_intercept = FIXED_INTERCEPT,
                                fdf = df_temp,
                                grouping_variable = "^Sample$")

## Add in degrees of freedom estimated from the randomized dataset
df_F_stats_out <- model_output_proper %>%
  dplyr::select(gene_name, H0_rss, H1_rss_sum, H0_df.residual, H1_df.residual_sum) %>% 
  unique() %>% 
  ungroup() %>% 
  left_join(dofs_from_randomized,
            by = "gene_name") %>% 
    ## Calculate F-statistic
  mutate(F_statistic = d2/d1 * (H0_rss - H1_rss_sum) / H1_rss_sum) %>% 
    ## Calculate P_value from F-statistic using the estimated degrees of freedom
  mutate(P_value = 1 - pf(F_statistic,
                    df1 = d1,
                    df2 = d2),
          ## Adjust P_value for multiple testing
         Adj_P_value = p.adjust(P_value,
                          method = "BH")) %>%
    ## Add in info on number of measured datapoints
  left_join(df_bins %>% 
              dplyr::select(gene_name, mean_n_datapoints_per_sample),
            by = "gene_name")



#### ---- Plots for checking, etc. ---  ####
plots[[length(plots)+1]] <-
  ggplot(df_F_stats_out,
         aes(x = F_statistic,
             y = after_stat(count))) +
    theme_classic() +
    theme(panel.background = element_rect(fill = "transparent")) +
    geom_histogram(bins = 1000)

plots[[length(plots)+1]] <-
  ggplot(df_F_stats_out,
         aes(x = log10(F_statistic),
             y = after_stat(count))) +
    theme_classic() +
    theme(panel.background = element_rect(fill = "transparent")) +
    geom_histogram(bins = 200)

plots[[length(plots)+1]] <-
  ggplot(df_F_stats_out,
         aes(x = P_value,
             y = after_stat(count))) +
    theme_classic() +
    theme(panel.background = element_rect(fill = "transparent")) +
    geom_histogram(bins = 1000)

plots[[length(plots)+1]] <-
  ggplot(df_F_stats_out,
         aes(x = P_value,
             y = mean_n_datapoints_per_sample)) +
    geom_point() +
    ggtitle("Mean number of datapoints per sample") +
    theme_classic() +
    theme(panel.background = element_rect(fill = "transparent"))

plots[[length(plots)+1]] <-
  ggplot(df_F_stats_out,
         aes(x = log10(H0_rss - H1_rss_sum),
             y = -log10(P_value),
             color = factor(bin))) +
    geom_point() +
    theme_classic() +
    theme(panel.background = element_rect(fill = "transparent")) +
    ggtitle("Bins with number of datapoints per sample")

plots[[length(plots)+1]] <-
  ggplot(df_F_stats_out,
         aes(x = log10(P_value),
             y = log10(Adj_P_value),
             color = factor(bin))) +
    geom_point() +
    theme_classic() +
    theme(panel.background = element_rect(fill = "transparent")) +
    geom_abline(linetype = "dashed")

## Example fits for representative purposes
plots[[length(plots)+1]] <-
  ggplot(df_temp %>% 
           filter(gene_name == "ARF4" | gene_name == "PSMD13" | gene_name == "ALDOA"),
         aes(x = time,
             y = new2old_tcc_adjusted,
             color = Sample)) +
    geom_point(size = 3) +
    geom_smooth(method = "lm",
                formula = y ~ x+0,
                se = FALSE) +
    theme_classic() +
    theme(panel.background = element_rect(fill = "transparent"),
          axis.text=element_text(size=10)) +
    ggtitle("H1") +
    scale_color_manual(values = COLOR_VECTOR_SAMPLE) +
    facet_wrap(~gene_name) +
  theme(panel.spacing = unit(2, "lines"))

plots[[length(plots)+1]] <-
  ggplot(df_temp %>% 
           filter(gene_name == "ARF4" | gene_name == "PSMD13" | gene_name == "ALDOA"),
         aes(x = time,
             y = new2old_tcc_adjusted)) +
    geom_point(size = 3) +
    geom_smooth(method = "lm",
                formula = y ~ x+0,
                se = FALSE) +
    theme_classic() +
    theme(panel.background = element_rect(fill = "transparent"),
          axis.text=element_text(size=10)) +
    ggtitle("H0") +
    facet_wrap(~gene_name) +
  theme(panel.spacing = unit(2, "lines"))

ggexport(ggarrange(plotlist = plots, nrow = 1, ncol = 1, widths = 12),
         filename = paste0("F-statistics_competing_fits_mouse_2DG_v_mouse_", VERSION, ".pdf"))

rm(df_bins, df_temp, plots, dofs_from_randomized, model_output_proper, model_output_randomized)
```

```{r, message = F, warning = F}
## Add in information, whether slower or faster
df_F_stats_mouse2DG_v_mouse <-
  df_fit_proteins_medians %>% 
  dplyr::select(-c(found_in_replicates, median_Thalf)) %>% 
  ## Remove negative kdegs
  filter(median_kdeg > 0 &
           Sample != "human") %>% 
  pivot_wider(names_from = Sample,
              values_from = median_kdeg) %>% 
  mutate(mouse2DG_vs_mouse = mouse_2DG - mouse) %>% 
  left_join(df_F_stats_out %>% 
              dplyr::select(gene_name, H0_rss, H1_rss_sum, F_statistic, Adj_P_value)) %>% 
  filter(!is.na(F_statistic)) %>% 
  ungroup() %>% 
  mutate(Hit_type = case_when(
    Adj_P_value <= ADJ_P_CUTOFF & mouse2DG_vs_mouse > 0 ~ "faster in 2DG",
    Adj_P_value <= ADJ_P_CUTOFF & mouse2DG_vs_mouse < 0 ~ "slower in 2DG",
    T ~ "no difference"), 
    Hit_type = factor(Hit_type,
                      levels = c("no difference",
                                 "faster in 2DG",
                                 "slower in 2DG")))
rm(df_F_stats_out)

## How many in each category?
df_F_stats_mouse2DG_v_mouse %>% 
  dplyr::select(Hit_type) %>% 
  table()
```

## Volcano-like plot from F statistics: mouse 2DG v mouse

```{r, message = F, warning = F}
FIG_WIDTH <- 3
FIG_HEIGHT <- 4

# PLOT_LIMITS <- c(-14, 0)

FILENAME <- "Figure3g_Volcano_mouse2DG_mouse"

ggplot(df_F_stats_mouse2DG_v_mouse,
       aes(x = sign(mouse2DG_vs_mouse)*sqrt(H0_rss - H1_rss_sum),
           y = log2(F_statistic + 1))) +
  geom_point(shape = 1,
             size = 0.8,
             stroke = 0.4,
             aes(color = Hit_type)) + 
  scale_color_manual(values = COLOR_HITS_2DG) +
  ylab(bquote(~ "log"["2"]*"(" * italic("F") *  "-statistic + 1)")) +
  xlab("effect size") +
  labs(color = "Hit") +
  theme_HH_scatter_paper +
  geom_text(data = tibble(
    x = 0.5,
    y = 1),
    aes(x,y, label = paste0("N = ", n_distinct(df_F_stats_mouse2DG_v_mouse$gene_name)))) + 
  theme(
    legend.position = "bottom",
    legend.direction = "vertical"
  )

ggsave(paste0(FILENAME, "_", VERSION, ".pdf"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

ggsave(paste0(FILENAME, "_", VERSION, ".svg"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

write.xlsx2(df_F_stats_mouse2DG_v_mouse %>% 
           as.data.frame() %>% 
             mutate(effect_size = sign(mouse2DG_vs_mouse)*sqrt(H0_rss - H1_rss_sum),
                    log2_Fstatistic_plus_one = log2(F_statistic + 1)),
         file = paste0(FILENAME, ".xlsx"),
         sheet = "Volcano",
         row.names = F)

```

## F statistic and abundance - mouse 2DG v mouse

```{r, message = F, warning = F}

df_plot <- df_F_stats_mouse2DG_v_mouse %>% 
         mutate(log2_mouse2DG_vs_mouse_Thalf = log2((log(2)/mouse_2DG)/(log(2)/mouse))) %>% 
  na.omit() %>% 
  left_join(
    df %>% 
      filter(Sample != "human") %>% 
      group_by(gene_name) %>% 
      summarise(median_abundance = log10(median(vsn_norm_TotalSignalSum, na.rm = T)))
  )

FIG_WIDTH <- 3
FIG_HEIGHT <- 3

FILENAME <- "FigureS7a_MA_plot_mouse2DG_v_mouse"

ggplot(df_plot,
       aes(x = median_abundance,
           y = log2_mouse2DG_vs_mouse_Thalf)) +
  geom_pointdensity(size = 0.5) +
  geom_hline(yintercept = 0,
             linewidth = 0.2) +
  scale_color_viridis() +
  ylab(label = bquote(~ "log" ["2"] * "(half life mouse 2DG / mouse)")) +
  xlab(label = bquote(~ "log" ["10"] * "(median protein abundance, a.u.)")) +
  theme_HH_scatter_paper +
  geom_text(data = tibble(
    x = 8,
    y = -3),
    aes(x,y, label = paste0("N = ", n_distinct(df_plot$gene_name)))) + 
  guides(color = FALSE)

ggsave(paste0(FILENAME, "_", VERSION, ".pdf"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

ggsave(paste0(FILENAME, "_", VERSION, ".svg"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

write.xlsx2(df_plot %>% 
           as.data.frame(),
         file = paste0(FILENAME, ".xlsx"),
         sheet = "MA_plot",
         row.names = F)

## Highlight hits from F statistic
FILENAME <- "FigureS7b_MA_plot_mouse2DG_v_mouse_with_Hits"

FIG_WIDTH <- 5
FIG_HEIGHT <- 3

ggplot(df_plot %>% 
         arrange(Hit_type),
       aes(x = median_abundance,
           y = log2_mouse2DG_vs_mouse_Thalf)) +
  geom_point(size = 0.5,
             aes(color = Hit_type)) +
  geom_hline(yintercept = 0,
             linewidth = 0.2) +
  scale_color_manual(values = COLOR_HITS_2DG) +
  ylab(label = bquote(~ "log" ["2"] * "(half life mouse 2DG / mouse)")) +
  xlab(label = bquote(~ "log" ["10"] * "(median protein abundance, a.u.)")) +
  theme_HH_scatter_paper +
  geom_text(data = tibble(
    x = 8,
    y = -3),
    aes(x,y, label = paste0("N = ", n_distinct(df_plot$gene_name)))) +
  labs(color = "Hit")

ggsave(paste0(FILENAME, "_", VERSION, ".pdf"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

ggsave(paste0(FILENAME, "_", VERSION, ".svg"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

write.xlsx2(df_plot %>% 
           as.data.frame(),
         file = paste0(FILENAME, ".xlsx"),
         sheet = "MA_plot",
         row.names = F)

```

# Difference in turnover and percentile turnover (human)

Use human protein turnover to rank proteins.

### Calculate rank order of turnover difference

```{r, message = F, warning = F}

## Take kdeg values from straight up fits
df_human_vs_mouse_ranks <- df_fit_proteins_medians %>% 
  filter(median_kdeg > 0 &
           Sample %in% c("mouse", "human")) %>% 
  dplyr::select(gene_name, Sample, median_Thalf) %>% 
  pivot_wider(names_from = Sample,
              values_from = median_Thalf) %>% 
  mutate(log2_human_vs_mouse = log2(human/mouse)) %>% 
  distinct() %>% 
  arrange(log2_human_vs_mouse)

df_mouse2DG_vs_mouse_ranks <- df_fit_proteins_medians %>% 
  filter(median_kdeg > 0 &
           Sample %in% c("mouse", "mouse_2DG")) %>%
  dplyr::select(gene_name, Sample, median_Thalf) %>% 
  pivot_wider(names_from = Sample,
              values_from = median_Thalf) %>% 
  mutate(log2_mouse2DG_vs_mouse = log2(mouse_2DG/mouse)) %>%
  distinct() %>% 
  arrange(log2_mouse2DG_vs_mouse)

df_human_vs_mouse2DG_ranks <- df_fit_proteins_medians %>% 
  filter(median_kdeg > 0 &
           Sample %in% c("human", "mouse_2DG")) %>%
  dplyr::select(gene_name, Sample, median_Thalf) %>% 
  pivot_wider(names_from = Sample,
              values_from = median_Thalf) %>% 
  mutate(log2_human_vs_mouse2DG = log2(human/mouse_2DG)) %>%
  distinct() %>% 
  arrange(log2_human_vs_mouse2DG)

```

human v mouse - fold change

```{r, message = F, warning = F}

FILENAME <- "Figure1e_deciles_human_v_mouse_fc"
FIG_WIDTH <- 2
FIG_HEIGHT <- 3

## Get percentile of kdeg in human
df_plot <- df_fit_proteins_medians %>%
  ungroup() %>% 
  filter(median_kdeg > 0 &
           Sample == "human") %>% 
  dplyr::select(gene_name, median_Thalf) %>% 
  na.omit() %>% 
  mutate(perc_rank = 1-percent_rank(median_Thalf),
         quantile = split_quantile(-median_Thalf, type = 10)) %>% 
  left_join(df_human_vs_mouse_ranks %>% 
              dplyr::select(gene_name, mouse, human)) %>% 
  mutate(quantile = case_when(
    quantile == "1" ~ "slowest",
    quantile == "10" ~ "fastest",
    T ~ quantile
  ) %>% 
    factor(levels = c("slowest", "2", "3", "4", "5", "6", "7", "8", "9", "fastest"))) %>% 
  ## Make signed log10 for difference
  mutate(signed_log10_diff_human_v_mouse = case_when(
    human - mouse > 0 ~ log10(human - mouse),
    human - mouse < 0 ~ -log10(abs(human-mouse))),
    human_v_mouse = log2(human/mouse)
  )

ggplot(df_plot,
       aes(x = quantile,
           y = human_v_mouse)) +
  geom_boxplot(lwd = LINEWIDTH,
               outlier.size = OUTLIER_SIZE,
               color = "grey20",
               fill = "grey20") +
  stat_summary(geom = "crossbar", width=0.65, lwd = 0.3, fatten = 0, color="white", fun.data = function(x){ return(c(y=median(x), ymin=median(x), ymax=median(x)))}) +
  theme_HH_paper +
  geom_hline(yintercept = 0,
             lwd = LINEWIDTH) +
  # ylab(label = bquote(~"log" [2] ~ "(T"["1/2, human"]~" / T"["1/2, mouse"]~")")) +
  ylab(label = bquote(~"log" [2] ~ "(half-life human / mouse")) +
  xlab("decile half-life\n(in human)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
ggsave(paste0(FILENAME, "", VERSION, ".pdf"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

ggsave(paste0(FILENAME, "", VERSION, ".svg"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

write.xlsx2(df_plot %>% 
           as.data.frame(),
         file = paste0(FILENAME, ".xlsx"),
         sheet = "all_data",
         row.names = F)

write.xlsx2(df_plot %>% 
           as.data.frame() %>% 
           group_by(quantile) %>% 
           summarise(median_signed_log10_diff_human_v_mouse = median(signed_log10_diff_human_v_mouse, na.rm = T),
                     median_diff_human_v_mouse = 10^median_signed_log10_diff_human_v_mouse) %>% 
           ungroup() %>% 
           as.data.frame(),
         file = paste0(FILENAME, ".xlsx"),
         sheet = "medians",
         row.names = F,
         append = TRUE)

```

mouse 2DG v mouse - fold change

```{r, message = F, warning = F}

FILENAME <- "Figure3h_deciles_mouse2DG_v_mouse_fc"
FIG_WIDTH <- 2
FIG_HEIGHT <- 3

## Get percentile of kdeg in human
df_plot <- df_fit_proteins_medians %>%
  ungroup() %>% 
  filter(median_kdeg > 0 &
           Sample == "human") %>% 
  dplyr::select(gene_name, median_Thalf) %>% 
  na.omit() %>% 
  mutate(perc_rank = 1-percent_rank(median_Thalf),
         quantile = split_quantile(-median_Thalf, type = 10)) %>% 
  left_join(df_mouse2DG_vs_mouse_ranks %>% 
              dplyr::select(gene_name, mouse, mouse_2DG)) %>% 
  mutate(quantile = case_when(
    quantile == "1" ~ "slowest",
    quantile == "10" ~ "fastest",
    T ~ quantile
  ) %>% 
    factor(levels = c("slowest", "2", "3", "4", "5", "6", "7", "8", "9", "fastest"))) %>% 
  ## Make signed log10 for difference
  mutate(signed_log10_diff_mouse2DG_v_mouse = case_when(
    mouse_2DG - mouse > 0 ~ log10(mouse_2DG - mouse),
    mouse_2DG - mouse < 0 ~ -log10(abs(mouse_2DG-mouse))),
    mouse2DG_v_mouse = log2(mouse_2DG/mouse)
  )

ggplot(df_plot,
       aes(x = quantile,
           y = mouse2DG_v_mouse)) +
  geom_boxplot(lwd = LINEWIDTH,
               outlier.size = OUTLIER_SIZE,
               color = "grey20",
               fill = "grey20") +
  stat_summary(geom = "crossbar", width=0.65, lwd = 0.3, fatten = 0, color="white", fun.data = function(x){ return(c(y=median(x), ymin=median(x), ymax=median(x)))}) +
  theme_HH_paper +
  geom_hline(yintercept = 0,
             lwd = LINEWIDTH) +
  # ylab(label = bquote(~"log" [2] ~ "(T"["1/2, mouse 2DG"]~" / T"["1/2, mouse"]~")")) +
  ylab(label = bquote(~"log" [2] ~ "(half-life mouse 2DG / mouse")) +
  xlab("decile half-life\n(in human)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
ggsave(paste0(FILENAME, "", VERSION, ".pdf"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

ggsave(paste0(FILENAME, "", VERSION, ".svg"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

write.xlsx2(df_plot %>% 
           as.data.frame(),
         file = paste0(FILENAME, ".xlsx"),
         sheet = "all_data",
         row.names = F)

write.xlsx2(df_plot %>% 
           as.data.frame() %>% 
           group_by(quantile) %>% 
           summarise(median_signed_log10_diff_mouse2DG_v_mouse = median(signed_log10_diff_mouse2DG_v_mouse, na.rm = T),
                     median_diff_mouse2DG_v_mouse = 10^signed_log10_diff_mouse2DG_v_mouse) %>% 
           ungroup() %>% 
           as.data.frame(),
         file = paste0(FILENAME, ".xlsx"),
         sheet = "medians",
         row.names = F,
         append = TRUE)

```
human v mouse 2DG, fold change

```{r, message = F, warning = F}

FILENAME <- "Deciles_human_v_mouse2DG_fc"
FIG_WIDTH <- 1.5
FIG_HEIGHT <- 3

## Get percentile of kdeg in human
df_plot <- df_fit_proteins_medians %>%
  ungroup() %>% 
  filter(median_kdeg > 0 &
           Sample == "human") %>% 
  dplyr::select(gene_name, median_Thalf) %>% 
  na.omit() %>% 
  mutate(perc_rank = 1-percent_rank(median_Thalf),
         quantile = split_quantile(-median_Thalf, type = 10)) %>% 
  left_join(df_human_vs_mouse2DG_ranks %>% 
              dplyr::select(gene_name, human, mouse_2DG)) %>% 
  mutate(quantile = case_when(
    quantile == "1" ~ "slowest",
    quantile == "10" ~ "fastest",
    T ~ quantile
  ) %>% 
    factor(levels = c("slowest", "2", "3", "4", "5", "6", "7", "8", "9", "fastest"))) %>% 
  ## Make signed log10 for difference
  mutate(signed_log10_diff_human_v_mouse2DG = case_when(
    human - mouse_2DG > 0 ~ log10(human - mouse_2DG),
    human - mouse_2DG < 0 ~ -log10(abs(human - mouse_2DG))),
    human_v_mouse2DG = log2(human/mouse_2DG)
  )

ggplot(df_plot,
       aes(x = quantile,
           y = human_v_mouse2DG)) +
  geom_boxplot(lwd = LINEWIDTH,
               outlier.size = OUTLIER_SIZE,
               color = "grey20",
               fill = "grey20") +
  stat_summary(geom = "crossbar", width=0.65, lwd = 0.3, fatten = 0, color="white", fun.data = function(x){ return(c(y=median(x), ymin=median(x), ymax=median(x)))}) +
  theme_HH_paper +
  geom_hline(yintercept = 0,
             lwd = LINEWIDTH) +
  # ylab(label = bquote(~"log" [2] ~ "(T"["1/2, human"]~" / T"["1/2, mouse 2DG"]~")")) +
  ylab(label = bquote(~"log" [2] ~ "(half-life human / mouse 2DG")) +
  xlab("decile half-life\n(in human)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
ggsave(paste0(FILENAME, "", VERSION, ".pdf"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

ggsave(paste0(FILENAME, "", VERSION, ".svg"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

write.xlsx2(df_plot %>% 
           as.data.frame(),
         file = paste0(FILENAME, ".xlsx"),
         sheet = "all_data",
         row.names = F)

write.xlsx2(df_plot %>% 
           as.data.frame() %>% 
           group_by(quantile) %>% 
           summarise(median_signed_log10_diff_human_v_mouse2DG = median(signed_log10_diff_human_v_mouse2DG, na.rm = T),
                     median_diff_human_v_mouse2DG = 10^signed_log10_diff_human_v_mouse2DG) %>% 
           ungroup() %>% 
           as.data.frame(),
         file = paste0(FILENAME, ".xlsx"),
         sheet = "medians",
         row.names = F,
         append = TRUE)

```

# GO enrichment analyses

## fgsea

Prepare GO term lists from Uniprot

```{r, message = F, warning = F}
# 
# ## Get GO terms from Uniprot
df_uniprot_GO <- read_delim("../221213_Uniprot.txt")

# df_uniprot_GO <- read_delim("https://rest.uniprot.org/uniprotkb/stream?fields=accession%2Cid%2Cgo_p%2Cgo_c%2Cgo_f%2Cgo_id%2Cgene_primary&format=tsv&query=%28human%29%20AND%20%28reviewed%3Atrue%29%20AND%20%28model_organism%3A9606%29", delim = "\t")
# 
# write.csv(df_uniprot_GO,
#           "df_uniprot_GO.txt",
#           row.names = F)
# 
# df_uniprot_GO <- read_csv("df_uniprot_GO.txt")

df_uniprot_GO <- df_uniprot_GO %>% 
  filter(!is.na(`Gene Ontology IDs`)) %>% 
  mutate(gene_name = map_chr(strsplit(`Gene Names`, " "), ~.x[1] %>% unlist())) %>% 
    dplyr::select(ID = Entry,
                  gene_name,
                  GO_CC = `Gene Ontology (cellular component)`,
                  GO_BP = `Gene Ontology (biological process)`,
                  GO_MF = `Gene Ontology (molecular function)`,
                  GO = `Gene Ontology IDs`)

df_uniprot_GO_CC <- df_uniprot_GO %>% 
  dplyr::select(gene_name, GO_CC) %>% 
  na.omit() %>% 
  mutate(GO_CC = strsplit(GO_CC, "; ")) %>% 
  unnest(GO_CC)

## Make list of sets
l_uniprot_GO_CC <- split(df_uniprot_GO_CC,
                       df_uniprot_GO_CC$GO_CC,
                       drop = TRUE) %>% 
  ## Choose first instance in each list (gene_name)
  lapply('[[', 1)

df_uniprot_GO_BP <- df_uniprot_GO %>% 
  dplyr::select(gene_name, GO_BP) %>% 
  na.omit() %>% 
  mutate(GO_BP = strsplit(GO_BP, "; ")) %>% 
  unnest(GO_BP)

l_uniprot_GO_BP <- split(df_uniprot_GO_BP,
                       df_uniprot_GO_BP$GO_BP,
                       drop = TRUE) %>% 
  ## Choose first instance in each list (gene_name)
  lapply('[[', 1)

df_uniprot_GO_MF <- df_uniprot_GO %>% 
  dplyr::select(gene_name, GO_MF) %>% 
  na.omit() %>% 
  mutate(GO_MF = strsplit(GO_MF, "; ")) %>% 
  unnest(GO_MF)

l_uniprot_GO_MF <- split(df_uniprot_GO_MF,
                       df_uniprot_GO_MF$GO_MF,
                       drop = TRUE) %>% 
  ## Choose first instance in each list (gene_name)
  lapply('[[', 1)

```

### human v mouse

```{r, message = F, warning = F, fig.height=10}

human_vs_mouse_ranks <- df_human_vs_mouse_ranks$log2_human_vs_mouse
names(human_vs_mouse_ranks) <- df_human_vs_mouse_ranks$gene_name

## Remove cases, which don't have a value in one of the species
human_vs_mouse_ranks <- na.omit(human_vs_mouse_ranks)

#############################

fgsea_CC <- fgsea(pathways = l_uniprot_GO_CC, 
                  stats    = human_vs_mouse_ranks,
                  minSize  = 15,
                  maxSize  = 500)

collapsedPathways <- collapsePathways(fgsea_CC[order(pval)][padj < 0.01], 
                                      l_uniprot_GO_CC, human_vs_mouse_ranks)
main_CC_Pathways <- fgsea_CC[pathway %in% collapsedPathways$mainPathways][
                         order(-NES), pathway]

plotGseaTable(l_uniprot_GO_CC[main_CC_Pathways], human_vs_mouse_ranks, fgsea_CC,
              gseaParam = 0.5)

# plotEnrichment(l_uniprot_GO_CC[["endoplasmic reticulum [GO:0005783]"]],
#                human_vs_mouse_ranks) + labs(title="endoplasmic reticulum [GO:0005783]")

#############################
fgsea_BP <- fgsea(pathways = l_uniprot_GO_BP, 
                  stats    = human_vs_mouse_ranks,
                  minSize  = 15,
                  maxSize  = 500)

collapsedPathways <- collapsePathways(fgsea_BP[order(pval)][padj < 0.01], 
                                      l_uniprot_GO_BP, human_vs_mouse_ranks)
main_BP_Pathways <- fgsea_BP[pathway %in% collapsedPathways$mainPathways][
                         order(-NES), pathway]

plotGseaTable(l_uniprot_GO_BP[main_BP_Pathways], human_vs_mouse_ranks, fgsea_BP,
              gseaParam = 0.5)

# plotEnrichment(l_uniprot_GO_BP[["gluconeogenesis [GO:0006094]"]],
#                human_vs_mouse_ranks) + labs(title="gluconeogenesis [GO:0006094]")

#############################
fgsea_MF <- fgsea(pathways = l_uniprot_GO_MF, 
                  stats    = human_vs_mouse_ranks,
                  minSize  = 15,
                  maxSize  = 500)

collapsedPathways <- collapsePathways(fgsea_MF[order(pval)][padj < 0.01], 
                                      l_uniprot_GO_MF, human_vs_mouse_ranks)
main_MF_Pathways <- fgsea_MF[pathway %in% collapsedPathways$mainPathways][
                         order(-NES), pathway]

plotGseaTable(l_uniprot_GO_MF[main_MF_Pathways], human_vs_mouse_ranks, fgsea_MF,
              gseaParam = 0.5)

# plotEnrichment(l_uniprot_GO_MF[["double-stranded RNA binding [GO:0003725]"]],
#                human_vs_mouse_ranks) + labs(title="double-stranded RNA binding [GO:0003725]")

```

```{r, message = F, warning = F}

fgsea_combined_human_v_mouse_main <- 
  rbind(
    fgsea_CC %>% 
      filter(pathway %in% main_CC_Pathways) %>% 
      mutate(GO_type = "CC"),
    fgsea_BP %>% 
      filter(pathway %in% main_BP_Pathways) %>% 
      mutate(GO_type = "BP"),
    fgsea_MF %>% 
      filter(pathway %in% main_MF_Pathways) %>% 
      mutate(GO_type = "MF")
  )

fgsea_combined_human_v_mouse <- 
  rbind(
    fgsea_CC %>% 
      filter(padj < 0.01) %>% 
      mutate(GO_type = "CC"),
    fgsea_BP %>%
      filter(padj < 0.01) %>% 
      mutate(GO_type = "BP"),
    fgsea_MF %>% 
      filter(padj < 0.01) %>%
      mutate(GO_type = "MF")
  )

fgsea_combined_human_v_mouse_all <- 
  rbind(
    fgsea_CC %>%  
      mutate(GO_type = "CC"),
    fgsea_BP %>%
      mutate(GO_type = "BP"),
    fgsea_MF %>% 
      mutate(GO_type = "MF")
  )

```

### mouse 2DG v mouse

```{r, message = F, warning = F, fig.height=10}

mouse2DG_vs_mouse_ranks <- df_mouse2DG_vs_mouse_ranks$log2_mouse2DG_vs_mouse
names(mouse2DG_vs_mouse_ranks) <- df_mouse2DG_vs_mouse_ranks$gene_name

mouse2DG_vs_mouse_ranks <- na.omit(mouse2DG_vs_mouse_ranks)
#############################

fgsea_CC <- fgsea(pathways = l_uniprot_GO_CC, 
                  stats    = mouse2DG_vs_mouse_ranks,
                  minSize  = 15,
                  maxSize  = 500)

collapsedPathways <- collapsePathways(fgsea_CC[order(pval)][padj < 0.01], 
                                      l_uniprot_GO_CC, mouse2DG_vs_mouse_ranks)
main_CC_Pathways <- fgsea_CC[pathway %in% collapsedPathways$mainPathways][
                         order(-NES), pathway]

plotGseaTable(l_uniprot_GO_CC[main_CC_Pathways], mouse2DG_vs_mouse_ranks, fgsea_CC,
              gseaParam = 0.5)

# plotEnrichment(l_uniprot_GO_CC[["endoplasmic reticulum [GO:0005783]"]],
#                mouse2DG_vs_mouse_ranks) + labs(title="endoplasmic reticulum [GO:0005783]")

#############################
fgsea_BP <- fgsea(pathways = l_uniprot_GO_BP, 
                  stats    = mouse2DG_vs_mouse_ranks,
                  minSize  = 15,
                  maxSize  = 500)

collapsedPathways <- collapsePathways(fgsea_BP[order(pval)][padj < 0.01], 
                                      l_uniprot_GO_BP, mouse2DG_vs_mouse_ranks)
main_BP_Pathways <- fgsea_BP[pathway %in% collapsedPathways$mainPathways][
                         order(-NES), pathway]

plotGseaTable(l_uniprot_GO_BP[main_BP_Pathways], mouse2DG_vs_mouse_ranks, fgsea_BP,
              gseaParam = 0.5)

# plotEnrichment(l_uniprot_GO_BP[["gluconeogenesis [GO:0006094]"]],
#                mouse2DG_vs_mouse_ranks) + labs(title="gluconeogenesis [GO:0006094]")

#############################
fgsea_MF <- fgsea(pathways = l_uniprot_GO_MF, 
                  stats    = mouse2DG_vs_mouse_ranks,
                  minSize  = 15,
                  maxSize  = 500)

collapsedPathways <- collapsePathways(fgsea_MF[order(pval)][padj < 0.01], 
                                      l_uniprot_GO_MF, mouse2DG_vs_mouse_ranks)
main_MF_Pathways <- fgsea_MF[pathway %in% collapsedPathways$mainPathways][
                         order(-NES), pathway]

plotGseaTable(l_uniprot_GO_MF[main_MF_Pathways], mouse2DG_vs_mouse_ranks, fgsea_MF,
              gseaParam = 0.5)

# plotEnrichment(l_uniprot_GO_MF[["double-stranded RNA binding [GO:0003725]"]],
#                mouse2DG_vs_mouse_ranks) + labs(title="double-stranded RNA binding [GO:0003725]")

```

```{r, message = F, warning = F}

fgsea_combined_mouse2DG_vs_mouse_main <- 
  rbind(
    fgsea_CC %>% 
      filter(pathway %in% main_CC_Pathways) %>% 
      mutate(GO_type = "CC"),
    fgsea_BP %>% 
      filter(pathway %in% main_BP_Pathways) %>% 
      mutate(GO_type = "BP"),
    fgsea_MF %>% 
      filter(pathway %in% main_MF_Pathways) %>% 
      mutate(GO_type = "MF")
  )

fgsea_combined_mouse2DG_vs_mouse <- 
  rbind(
    fgsea_CC %>% 
      filter(padj < 0.01) %>% 
      mutate(GO_type = "CC"),
    fgsea_BP %>%
      filter(padj < 0.01) %>% 
      mutate(GO_type = "BP"),
    fgsea_MF %>% 
      filter(padj < 0.01) %>%
      mutate(GO_type = "MF")
  )

fgsea_combined_mouse2DG_vs_mouse_all <- 
  rbind(
    fgsea_CC %>% 
      mutate(GO_type = "CC"),
    fgsea_BP %>%
      mutate(GO_type = "BP"),
    fgsea_MF %>% 
      mutate(GO_type = "MF")
  )

```

### human vs mouse 2DG

```{r, message = F, warning = F}

human_vs_mouse2DG_ranks <- df_human_vs_mouse2DG_ranks$log2_human_vs_mouse2DG
names(human_vs_mouse2DG_ranks) <- df_human_vs_mouse2DG_ranks$gene_name

human_vs_mouse2DG_ranks <- na.omit(human_vs_mouse2DG_ranks)

#############################

fgsea_CC <- fgsea(pathways = l_uniprot_GO_CC, 
                  stats    = human_vs_mouse2DG_ranks,
                  minSize  = 15,
                  maxSize  = 500)

collapsedPathways <- collapsePathways(fgsea_CC[order(pval)][padj < 0.01], 
                                      l_uniprot_GO_CC, human_vs_mouse2DG_ranks)
main_CC_Pathways <- fgsea_CC[pathway %in% collapsedPathways$mainPathways][
                         order(-NES), pathway]

plotGseaTable(l_uniprot_GO_CC[main_CC_Pathways], human_vs_mouse2DG_ranks, fgsea_CC,
              gseaParam = 0.5)

# plotEnrichment(l_uniprot_GO_CC[["endoplasmic reticulum [GO:0005783]"]],
#                human_vs_mouse2DG_ranks) + labs(title="endoplasmic reticulum [GO:0005783]")

#############################
fgsea_BP <- fgsea(pathways = l_uniprot_GO_BP, 
                  stats    = human_vs_mouse2DG_ranks,
                  minSize  = 15,
                  maxSize  = 500)

collapsedPathways <- collapsePathways(fgsea_BP[order(pval)][padj < 0.01], 
                                      l_uniprot_GO_BP, human_vs_mouse2DG_ranks)
main_BP_Pathways <- fgsea_BP[pathway %in% collapsedPathways$mainPathways][
                         order(-NES), pathway]

plotGseaTable(l_uniprot_GO_BP[main_BP_Pathways], human_vs_mouse2DG_ranks, fgsea_BP,
              gseaParam = 0.5)

# plotEnrichment(l_uniprot_GO_BP[["gluconeogenesis [GO:0006094]"]],
#                human_vs_mouse2DG_ranks) + labs(title="gluconeogenesis [GO:0006094]")

#############################
fgsea_MF <- fgsea(pathways = l_uniprot_GO_MF, 
                  stats    = human_vs_mouse2DG_ranks,
                  minSize  = 15,
                  maxSize  = 500)

collapsedPathways <- collapsePathways(fgsea_MF[order(pval)][padj < 0.01], 
                                      l_uniprot_GO_MF, human_vs_mouse2DG_ranks)
main_MF_Pathways <- fgsea_MF[pathway %in% collapsedPathways$mainPathways][
                         order(-NES), pathway]

plotGseaTable(l_uniprot_GO_MF[main_MF_Pathways], human_vs_mouse2DG_ranks, fgsea_MF,
              gseaParam = 0.5)

# plotEnrichment(l_uniprot_GO_MF[["double-stranded RNA binding [GO:0003725]"]],
#                human_vs_mouse2DG_ranks) + labs(title="double-stranded RNA binding [GO:0003725]")

```

```{r, message = F, warning = F}

fgsea_combined_human_vs_mouse2DG_main <- 
  rbind(
    fgsea_CC %>% 
      filter(pathway %in% main_CC_Pathways) %>% 
      mutate(GO_type = "CC"),
    fgsea_BP %>% 
      filter(pathway %in% main_BP_Pathways) %>% 
      mutate(GO_type = "BP"),
    fgsea_MF %>% 
      filter(pathway %in% main_MF_Pathways) %>% 
      mutate(GO_type = "MF")
  )

fgsea_combined_human_vs_mouse2DG <- 
  rbind(
    fgsea_CC %>% 
      filter(padj < 0.01) %>% 
      mutate(GO_type = "CC"),
    fgsea_BP %>%
      filter(padj < 0.01) %>% 
      mutate(GO_type = "BP"),
    fgsea_MF %>% 
      filter(padj < 0.01) %>%
      mutate(GO_type = "MF")
  )

fgsea_combined_human_vs_mouse2DG_all <- 
  rbind(
    fgsea_CC %>%  
      mutate(GO_type = "CC"),
    fgsea_BP %>% 
      mutate(GO_type = "BP"),
    fgsea_MF %>%
      mutate(GO_type = "MF")
  )

```

```{r, message = F, warning = F}
rm(fgsea_CC, fgsea_BP, fgsea_MF, main_BP_Pathways, main_CC_Pathways, main_MF_Pathways)
```

## Dotplots of all enrichments together

```{r, fig.height=3, fig.width=5}

# df_main_pathways <-
#   rbind(fgsea_combined_human_v_mouse_main %>% 
#           mutate(comparison = "human v mouse"),
#         fgsea_combined_human_vs_mouse2DG_main %>% 
#           mutate(comparison = "human v mouse 2DG"),
#         fgsea_combined_mouse2DG_vs_mouse_main %>% 
#           mutate(comparison = "mouse 2DG v mouse"))
# 
# 
# FILENAME <- "All_GO_enrichments_main_pathways"
# FIG_WIDTH <- 5
# FIG_HEIGHT <- 3.25
# 
# plots <- list()
# 
# ## cycle through each pathway to get ordering correct
# for(i in unique(df_main_pathways$comparison)){
#   df_temp <- df_main_pathways %>% 
#     filter(comparison == i)
#   for(j in unique(df_temp$GO_type)){
#     df_temp1 <- df_temp %>% 
#       filter(GO_type == j)
#     
#     if(nrow(df_temp1)>0){
#       
#       plots[[length(plots)+1]] <-
#         ## Dotplot of main pathways
#         ggplot(df_temp1 %>% 
#                  group_by(comparison, GO_type) %>% 
#                  mutate(pathway = fct_reorder(pathway, NES)),
#                aes(x = NES,
#                    y = pathway,
#                    size = size,
#                    fill = -log10(padj))) +
#         geom_point(shape = 21,
#                    linewidth = LINEWIDTH,
#                    color = "grey20") +
#         theme_HH_paper +
#         geom_vline(xintercept = 0,
#                    lwd = LINEWIDTH,
#                    linetype = "dashed",
#                    color = "grey") +
#         xlim(c(-2.8,2.8))+
#         scale_fill_gradient(
#           low = "grey90",
#           high = "firebrick",
#           limits = c(-log10(0.05), 6),
#           oob = squish
#         ) +
#         labs(fill = bquote(-log[10]*"(adj. P val)")) +
#         ylab("") +
#         theme(legend.key.size = unit(0.3, "cm")) +
#         scale_y_discrete(labels = function(x) str_wrap(x, width = 40)) +
#         facet_wrap(~comparison + GO_type,
#                    scales = "free")
#       
#     }
#   }
# }
# 
# ggexport(ggarrange(plotlist = plots, nrow = 1, ncol = 1, widths = 12),
#          filename = paste0(FILENAME, "_", VERSION,".pdf"),
#          width = FIG_WIDTH, height = FIG_HEIGHT)

#################################

df_all_pathways <-
  rbind(fgsea_combined_human_v_mouse %>% 
          filter(padj < 0.01) %>% 
          mutate(comparison = "human v mouse"),
        fgsea_combined_human_vs_mouse2DG %>% 
          filter(padj < 0.01) %>% 
          mutate(comparison = "human v mouse 2DG"),
        fgsea_combined_mouse2DG_vs_mouse %>% 
          filter(padj < 0.01) %>% 
          mutate(comparison = "mouse 2DG v mouse"))


FILENAME <- "All_GO_enrichments_all_pathways"
FIG_WIDTH <- 5
FIG_HEIGHT <- 3.25

plots <- list()

## cycle through each pathway to get ordering correct
for(i in unique(df_all_pathways$comparison)){
  df_temp <- df_all_pathways %>% 
    filter(comparison == i)
  for(j in unique(df_temp$GO_type)){
    df_temp1 <- df_temp %>% 
      filter(GO_type == j)
    
    if(nrow(df_temp1)>0){
      
      plots[[length(plots)+1]] <-
        ## Dotplot of main pathways
        ggplot(df_temp1 %>% 
                 group_by(comparison, GO_type) %>% 
                 mutate(pathway = fct_reorder(pathway, NES)),
               aes(x = NES,
                   y = pathway,
                   size = size,
                   fill = -log10(padj))) +
        geom_point(shape = 21,
                   linewidth = LINEWIDTH,
                   color = "grey20") +
        theme_HH_paper +
        geom_vline(xintercept = 0,
                   lwd = LINEWIDTH,
                   linetype = "dashed",
                   color = "grey") +
        xlim(c(-2.8,2.8))+
        scale_fill_gradient(
          low = "grey90",
          high = "firebrick",
          limits = c(-log10(0.05), 6),
          oob = squish
        ) +
        labs(fill = bquote(-log[10]*"(adj. P val)")) +
        ylab("") +
        theme(legend.key.size = unit(0.3, "cm")) +
        scale_y_discrete(labels = function(x) str_wrap(x, width = 40)) +
        facet_wrap(~comparison + GO_type,
                   scales = "free")
      
    }
  }
}

ggexport(ggarrange(plotlist = plots, nrow = 1, ncol = 1, widths = 12),
         filename = paste0(FILENAME, "_", VERSION, ".pdf"),
         width = FIG_WIDTH, height = FIG_HEIGHT)

```

### Export GO enrichments

```{r}

all_fgsea <-
  rbind(
    fgsea_combined_human_v_mouse_all %>% 
      mutate(comparison = "human v mouse"),
    fgsea_combined_human_vs_mouse2DG_all %>% 
      mutate(comparison = "human v mouse2DG"),
    fgsea_combined_mouse2DG_vs_mouse_all %>% 
      mutate(comparison = "mouse2DG v mouse")
  ) %>% 
  rowwise() %>% 
  mutate(leadingEdge = paste(leadingEdge, collapse = ", "))

write.csv(all_fgsea,
          "Gene_Set_Enrichment_results_fgsea.csv",
          row.names = F)



```

# Subcellular localization

Get unique and non-unique annotation

```{r}

## Only look at the following GO terms
GO_id <- tibble(
  "GO" =
    c("GO:0005576",
      "GO:0031012",
      "GO:0005886",
      "GO:0005856",
      "GO:0005737",
      "GO:0005829",
      "GO:0005739",
      "GO:0005777",
      "GO:0005764",
      "GO:0005783",
      "GO:0005794",
      "GO:0005634",
      "GO:0031965",
      "GO:0005654",
      "GO:0005730"),
  "CC" =
    c("extracellular region",
      "extracellular matrix",
      "plasma membrane",
      "cytoskeleton",
      "cytoplasm",
      "cytosol",
      "mitochondrion",
      "peroxisome",
      "lysosome",
      "ER",
      "Golgi apparatus",
      "nucleus",
      "nuclear membrane",
      "nucleoplasm",
      "nucleolus"))

  df_in <- df_uniprot_GO %>% 
    dplyr::select(ID, gene_name, GO) %>% 
    mutate(GO = strsplit(GO, "; ")) %>% 
    unnest(GO) %>% 
    left_join(GO_id) %>% 
    group_by(ID, gene_name) %>% 
    mutate(n_CC = n_distinct(CC, na.rm = T))
  
  ## Make lists of all annotations (i.e. non-unique and unique)
  df_all <- df_in %>% 
    dplyr::select(ID, gene_name, CC) %>% 
    na.omit() %>% 
    distinct() %>% 
    group_by(ID, gene_name) %>% 
    summarise(CC_all = paste(CC, collapse = ";"))
  
  ## Combine everything
  df_CC_annotation <- df_in %>% 
    dplyr::select(ID, gene_name) %>% 
    distinct() %>% 
    left_join(
      df_all
    )

rm(df_all, df_in)
```

## H v M - Non-unique annotation

```{r}

FILENAME <- "Figure2a_CC_annotation_human_v_mouse_line_at_zero"
FIG_WIDTH <- 4
FIG_HEIGHT <- 4

df_nonunique_annotation <- df_CC_annotation %>% 
  dplyr::select(gene_name, CC_all) %>% 
  mutate(CC_all = strsplit(CC_all, ";")) %>% 
  unnest(CC_all) %>% 
  na.omit()

df_plot <- 
  rbind(df_human_vs_mouse_ranks %>% 
          mutate(CC_all = "all"),
        df_human_vs_mouse_ranks %>% 
          left_join(df_nonunique_annotation) %>% 
          na.omit()
  ) %>% 
  mutate(CC_all = factor(CC_all, levels = c(
    "all", "extracellular matrix", "extracellular region", "plasma membrane", "cytoskeleton", "cytoplasm", "cytosol", "mitochondrion", "lysosome", "Golgi apparatus", "peroxisome", "ER", "nuclear membrane", "nucleus", "nucleoplasm", "nucleolus"
  )))

ggplot(df_plot,
       aes(x = CC_all,
           y = log2_human_vs_mouse,
           fill = CC_all,
           color = CC_all)) +
    geom_boxplot(lwd = LINEWIDTH,
               outlier.size = OUTLIER_SIZE) +
  stat_summary(geom = "crossbar", width=0.65, lwd = 0.3, fatten = 0, color="white", fun.data = function(x){ return(c(y=median(x), ymin=median(x), ymax=median(x)))}) +
  theme_HH_paper +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  stat_summary(fun.data = function(x){
    data.frame(y=min(x)-0.5,
               label = length(x))
  }, geom="text", color = "black",
  size = 2) +
  ggtitle("GO: CC annotation") +
  # geom_hline(yintercept = df_plot %>% filter(CC_all == "all") %>% pull(log2_human_vs_mouse) %>% median(na.rm = T),
  #            lwd = LINEWIDTH) +
  geom_hline(yintercept = 0,
             lwd = LINEWIDTH) +  
  # stat_compare_means(
  #   label = "p.signif",
  #   method = "t.test", ref.group = "all",
  #   hide.ns = T
  # ) +
  scale_fill_manual(values = c(COLOR_CC,"all" = "grey")) +
  scale_color_manual(values = c(COLOR_CC,"all" = "grey")) +
    theme(legend.position = "none") +
  xlab("") +
  # ylab(label = bquote(~ "T" [1/2] ~ "log" ["2"] * "(human / mouse)"))
  ylab(label = bquote(~ "log" ["2"] * "(half life human / mouse)"))


ggsave(paste0(FILENAME, "", VERSION, ".pdf"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

ggsave(paste0(FILENAME, "", VERSION, ".svg"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

write.xlsx2(df_plot %>% 
           as.data.frame(),
         file = paste0(FILENAME, ".xlsx"),
         sheet = "all_data",
         row.names = F)

write.xlsx2(df_plot %>% 
           as.data.frame() %>% 
           group_by(CC_all) %>% 
           summarise(median_log2_human_vs_mouse = median(log2_human_vs_mouse, na.rm = T)) %>% 
           ungroup() %>% 
           as.data.frame(),
         file = paste0(FILENAME, ".xlsx"),
         sheet = "medians",
         row.names = F,
         append = TRUE)

write.xlsx2(compare_means(log2_human_vs_mouse ~ CC_all, data = df_plot, group_by = "CC_all", method = "t.test", ref.group = "all") %>% 
           as.data.frame() %>% 
             ## Add significance levels from adjusted p values
             mutate(p.adj.signif = case_when(
               p.adj < 0.0001 ~ "****",
               p.adj < 0.001 ~ "***",
               p.adj < 0.01 ~ "**",
               p.adj < 0.05 ~ "*",
               T ~ "ns"
             )),
         file = paste0(FILENAME, ".xlsx"),
         sheet = "t_test",
         row.names = F,
         append = TRUE)

```

## H v M2DG - Non-unique annotation

```{r}

FILENAME <- "Figure4c_CC_annotation_human_v_mouse2DG_line_at_zero"
FIG_WIDTH <- 4
FIG_HEIGHT <- 4

df_nonunique_annotation <- df_CC_annotation %>% 
  dplyr::select(gene_name, CC_all) %>% 
  mutate(CC_all = strsplit(CC_all, ";")) %>% 
  unnest(CC_all) %>% 
  na.omit()

df_plot <- 
  rbind(df_human_vs_mouse2DG_ranks %>% 
          mutate(CC_all = "all"),
        df_human_vs_mouse2DG_ranks %>% 
          left_join(df_nonunique_annotation) %>% 
          na.omit()
  ) %>% 
  mutate(CC_all = factor(CC_all, levels = c(
    "all", "extracellular matrix", "extracellular region", "plasma membrane", "cytoskeleton", "cytoplasm", "cytosol", "mitochondrion", "lysosome", "Golgi apparatus", "peroxisome", "ER", "nuclear membrane", "nucleus", "nucleoplasm", "nucleolus"
  )))

ggplot(df_plot,
       aes(x = CC_all,
           y = log2_human_vs_mouse2DG,
           fill = CC_all,
           color = CC_all)) +
    geom_boxplot(lwd = LINEWIDTH,
               outlier.size = OUTLIER_SIZE) +
  stat_summary(geom = "crossbar", width=0.65, lwd = 0.3, fatten = 0, color="white", fun.data = function(x){ return(c(y=median(x), ymin=median(x), ymax=median(x)))}) +
  theme_HH_paper +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  stat_summary(fun.data = function(x){
    data.frame(y=min(x)-0.5,
               label = length(x))
  }, geom="text", color = "black",
  size = 2) +
  ggtitle("GO: CC annotation") +
  # geom_hline(yintercept = df_plot %>% filter(CC_all == "all") %>% pull(log2_human_vs_mouse2DG) %>% median(na.rm = T),
  #            lwd = LINEWIDTH) +
  geom_hline(yintercept = 0,
             lwd = LINEWIDTH) +  
  # stat_compare_means(
  #   label = "p.signif",
  #   method = "t.test", ref.group = "all",
  #   hide.ns = T
  # ) +
  scale_fill_manual(values = c(COLOR_CC,"all" = "grey")) +
  scale_color_manual(values = c(COLOR_CC,"all" = "grey")) +
    theme(legend.position = "none") +
  xlab("") +
  # ylab(label = bquote(~ "T" [1/2] ~ "log" ["2"] * "(human / mouse - 2DG)"))
  ylab(label = bquote(~ "log" ["2"] * "(half life human / mouse_2DG)"))

ggsave(paste0(FILENAME, "", VERSION, ".pdf"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

ggsave(paste0(FILENAME, "", VERSION, ".svg"),
       width = FIG_WIDTH,
       height = FIG_HEIGHT)

write.xlsx2(df_plot %>% 
           as.data.frame(),
         file = paste0(FILENAME, ".xlsx"),
         sheet = "all_data",
         row.names = F)

write.xlsx2(df_plot %>% 
           as.data.frame() %>% 
           group_by(CC_all) %>% 
           summarise(median_log2_human_vs_mouse2DG = median(log2_human_vs_mouse2DG, na.rm = T)) %>% 
           ungroup() %>% 
           as.data.frame(),
         file = paste0(FILENAME, ".xlsx"),
         sheet = "medians",
         row.names = F,
         append = TRUE)

write.xlsx2(compare_means(log2_human_vs_mouse2DG ~ CC_all, data = df_plot, group_by = "CC_all", method = "t.test", ref.group = "all") %>% 
           as.data.frame() %>% 
             ## Add significance levels from adjusted p values
             mutate(p.adj.signif = case_when(
               p.adj < 0.0001 ~ "****",
               p.adj < 0.001 ~ "***",
               p.adj < 0.01 ~ "**",
               p.adj < 0.05 ~ "*",
               T ~ "ns"
             )),
         file = paste0(FILENAME, ".xlsx"),
         sheet = "t_test",
         row.names = F,
         append = TRUE)

```

## Export all results

```{r}

write.csv(df_fit_proteins_medians,
          "median_Half-lives_over_replicates.csv",
          row.names = F)

df_export <-
  rbind(
    df_F_stats_human_v_mouse %>%
          dplyr::select(gene_name, Adj_P_value,  Hit_type) %>%
          left_join(df_human_vs_mouse_ranks %>% 
                      dplyr::select(gene_name, log2_human_vs_mouse)) %>%
          mutate(comparison = "human v mouse") %>%
          dplyr::rename(logFC = log2_human_vs_mouse),
        df_human_vs_mouse2DG_ranks %>% 
          dplyr::select(gene_name, logFC = log2_human_vs_mouse2DG) %>% 
          mutate(comparison = "human v mouse 2DG",
                 Adj_P_value = NA,
                 Hit_type = NA),
        df_F_stats_mouse2DG_v_mouse %>%
          dplyr::select(gene_name, Adj_P_value, Hit_type) %>% 
          left_join(df_mouse2DG_vs_mouse_ranks %>% 
                      dplyr::select(gene_name, log2_mouse2DG_vs_mouse)) %>%
          mutate(comparison = "mouse 2DG v mouse") %>% 
          dplyr::rename(logFC = log2_mouse2DG_vs_mouse)) %>% 
  ## Add annotations
  left_join(df_uniprot_GO %>% 
              dplyr::select(gene_name, GO_CC, GO_BP, GO_MF))

write.csv(df_export,
          "Fold_changes_and_annotations.csv",
          row.names = F)

rm(df_export, df_export_kdegs)
```

## Placeholder

```{r}

```

